---
title: "HFTV1_REANNOTATION"
author: "mfisada"
date: "2026-02-05"
output: html_document
---

# === PACKAGES ===============================================================
```{r}
library(rtracklayer)
library(tidyverse)
library(tidyr)
library(ivs) # For overlaps
library(data.table) # For conversion from df to df
library(Biostrings)
library(phylotools) # For reading FASTA files
library(seqinr) # For exporting FASTA files
library(plyr)
library(dplyr)
library(plyranges) # For reading bigwigs
```

# === Hg/HFTV1 | IMPORT NCBI ANNOTATIONS =====================================
```{r}
# Create GRanges objects from annotation files
HFTV1_NCBI_gtf_gr <- 
  rtracklayer::import(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
      "GCF_004208775.1/genomic.gtf", sep = ""))
Hg_NCBI_gtf_gr <- 
  rtracklayer::import(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/H_GIBBONSII/",
      "DL_20250521/GCF_014969745.1/Hg_genomic.gtf", sep = ""))

# Convert GRanges to data.frame
HFTV1_NCBI_gtf_df <- as.data.frame(HFTV1_NCBI_gtf_gr)
Hg_NCBI_gtf_df <- as.data.frame(Hg_NCBI_gtf_gr)


# Reformat NCBI GTF to meet formatting standards
# https://agat.readthedocs.io/en/latest/gxf.html
# https://biocorecrg.github.io/PhD_course_genomics_format_2021/gtf_format.html
# https://www.biobam.com/differences-between-gtf-and-gff-files-in-genomic-data-analysis/
# Column order:
# col1    seqname
# col2    source
# col3    feature
# col4    start
# col5    end
# col6    score
# col7    strand
# col8    frame
# col9    attribute

HFTV1_NCBI_gtf_df <-
  HFTV1_NCBI_gtf_df %>%
  # Remove columns
  select(!width) %>%
  # Rename columns
  dplyr::rename(
    seqname = seqnames,
    feature = type,
    frame = phase) %>%
  # Relocate columns
  relocate(source, .after = seqname) %>%
  relocate(feature, .after = source) %>%
  relocate(start, .after = feature) %>%
  relocate(end, .after = start) %>%
  relocate(score, .after = end) %>%
  relocate(strand, .after = score) %>%
  relocate(frame, .after = strand)

Hg_NCBI_gtf_df <-
  Hg_NCBI_gtf_df %>%
  # Remove columns
  select(!width) %>%
  # Rename columns
  dplyr::rename(
    seqname = seqnames,
    feature = type,
    frame = phase) %>%
  # Relocate columns
  relocate(source, .after = seqname) %>%
  relocate(feature, .after = source) %>%
  relocate(start, .after = feature) %>%
  relocate(end, .after = start) %>%
  relocate(score, .after = end) %>%
  relocate(strand, .after = score) %>%
  relocate(frame, .after = strand)
```

# === HFTV1 | ANNOTATION REVISIONS FROM CRYO-EM PAPER ========================
```{r}
# Annotation information from: Daniel X. Zhang et al. ,Cryo-EM resolves the
# structure of the archaeal dsDNA virus HFTV1 from head to tail.Sci.
# Adv.11,eadx1178(2025).DOI:10.1126/sciadv.adx1178

# Revise pre-existing HFTV1 annotations

# GTF
HFTV1_NCBI_gtf_df <- 
  HFTV1_NCBI_gtf_df %>%
  mutate(
    # Change annotated product based on conditions below
    product = 
      case_when(
        
        # Change annotated product for individual locus tags. Exclude rows
        # that have no predicted product; this is the case for the "gene"
        # entry of each feature, which by default has no predicted product.
        
        # Structural components from Zhang et al., 2024, unpublished.
        locus_tag == "M1M17_gp13" & 
          !is.na(product) ~ "portal protein",
        locus_tag == "M1M17_gp14" & 
          !is.na(product) ~ "TMP-DNA spacer",
        locus_tag == "M1M17_gp18" & 
          !is.na(product) ~ "capsid stabilization protein",
        locus_tag == "M1M17_gp19" & 
          !is.na(product) ~ "major capsid protein",
        locus_tag == "M1M17_gp21" & 
          !is.na(product) ~ "portal interface protein",
        locus_tag == "M1M17_gp26" & 
          !is.na(product) ~ "head-tail connector",
        locus_tag == "M1M17_gp29" & 
          !is.na(product) ~ "tail completion protein",
        locus_tag == "M1M17_gp30" & 
          !is.na(product) ~ "turret base protein",
        locus_tag == "M1M17_gp31" & 
          !is.na(product) ~ "turret head protein",
        locus_tag == "M1M17_gp35" & 
          !is.na(product) ~ "tail tube protein",
        locus_tag == "M1M17_gp40" & 
          !is.na(product) ~ "tape measure protein",
        locus_tag == "M1M17_gp41" & 
          !is.na(product) ~ "baseplate-tail connector",
        locus_tag == "M1M17_gp42" & 
          !is.na(product) ~ "tail fibre protein",
        locus_tag == "M1M17_gp44" & 
          !is.na(product) ~ "baseplate hub",
        
        # Set original product values as default if none of the conditions
        # above are met. 
        TRUE ~ product))

HFTV1_NCBI_gtf_gr <-
  makeGRangesFromDataFrame(
    HFTV1_NCBI_gtf_df, keep.extra.columns = TRUE)

# === EXPORT ===

# export(
#   HFTV1_NCBI_gtf_gr,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
#     "GCF_004208775.1/ANNOT_FUNCTION_REVISION/HFTV1_NCBI_CRYOEM.gtf",
#     sep = ""),
#   "gff3")
# 
# write_csv(
#   HFTV1_NCBI_gtf_df,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
#     "GCF_004208775.1/ANNOT_FUNCTION_REVISION/HFTV1_NCBI_CRYOEM.csv",
#     sep = ""))

```

# === HFTV1 | IMPORT PHAROKKA/PHOLD (P/P) ANNOTATIONS ========================
```{r}
# Create GRanges object from annotation file
HFTV1_PP_gff_gr <- 
  rtracklayer::import(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
      "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/a_PHOLD_OUTPUT/",
      "phold_no_dna.gff3", 
      sep = ""))
# Manually removed entry in phold output file corresponding to full genome. 

# Convert GRanges to data.frame
HFTV1_PP_gff_df <- 
  as.data.frame(HFTV1_PP_gff_gr) %>%
  mutate(
    # All Pharokka-predicted features are assigned a phase value of 0 in the
    # Pharokka GFF, but they are then switched to a phase value of 1 after
    # structural homology predictions by Phold. This is presumably a technical
    # error of the Phold GFF reformatting/output process. Consequently, convert
    # all phase values from 1 to 0. 
    phase = 0,
    # Set all scores to NA.
    score = NA)

# Lines for report later
HFTV1_PP_ct_line <-
  paste(
    "Number of PP features: ",
    nrow(HFTV1_PP_gff_df), sep = "")

HFTV1_PP_CDS_line <-
  paste(
    "Number of PP feature type 'CDS': ",
    nrow(
      HFTV1_PP_gff_df %>% filter(type == "CDS")), sep = "")
HFTV1_PP_tRNA_line <-
  paste(
    "Number of PP feature type 'tRNA': ",
    nrow(
      HFTV1_PP_gff_df %>% filter(type == "tRNA")), sep = "")
```

# === HFTV1 | OVERLAPS BETWEEN P/P AND NCBI FEATURES =========================
```{r}

# === Subset HFTV1 GFF to compare with P/P features ===

# Subset to CDSs (the only feature type in NCBI annotation besides gene,
# start_codon, and stop_codon).
HFTV1_NCBI_CDS_df <-
  HFTV1_NCBI_gtf_df %>%
  filter(feature == "CDS")

HFTV1_NCBI_CDS_ct_line <- 
  paste(
    "Number of NCBI CDSs: ",
    nrow(HFTV1_NCBI_CDS_df), sep = "")

# Convert to GRanges to assess overlaps
HFTV1_NCBI_CDS_gr <- 
  makeGRangesFromDataFrame(HFTV1_NCBI_CDS_df, keep.extra.columns = TRUE)


# === Find overlaps between NCBI and P/P-predicted features ===

# P/P as query
HFTV1_PP_NCBI_overlaps <-
  findOverlaps(
    HFTV1_PP_gff_gr,
    HFTV1_NCBI_CDS_gr,
    minoverlap = 60)
HFTV1_PP_NCBI_hits_df <-
  data.frame(
    HFTV1_PP_gff_gr[queryHits(HFTV1_PP_NCBI_overlaps),],
    HFTV1_NCBI_CDS_gr[subjectHits(HFTV1_PP_NCBI_overlaps),])

# NCBI as query
HFTV1_NCBI_PP_overlaps <-
  findOverlaps(
    HFTV1_NCBI_CDS_gr,
    HFTV1_PP_gff_gr,
    minoverlap = 60)
HFTV1_NCBI_PP_hits_df <-
  data.frame(
    HFTV1_NCBI_CDS_gr[queryHits(HFTV1_NCBI_PP_overlaps),],
    HFTV1_PP_gff_gr[subjectHits(HFTV1_NCBI_PP_overlaps),])


# === Overlapping features ===

# Output from both approaches below should be similar. 

# P/P as query
HFTV1_PP_NCBI_overlap_df <- 
  inner_join(
    HFTV1_PP_gff_df, 
    HFTV1_PP_NCBI_hits_df[,c("start", "end")], 
    by = c("start", "end"))

HFTV1_PP_NCBI_overlap_line <- 
  paste(
    "Number of distinct overlapping features (P/P as query): ",
    n_distinct(HFTV1_PP_NCBI_overlap_df$ID), sep = "")

HFTV1_PP_NCBI_overlap_gr <- 
  makeGRangesFromDataFrame(HFTV1_PP_NCBI_overlap_df, keep.extra.columns = TRUE)

# export(
#   HFTV1_PP_NCBI_overlap_gr,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
#     "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
#     "b_PP_NCBI_OVERLAPS/HFTV1_PP_NCBI_overlap.gff", sep = ""),
#   "gff3")
# write_csv(
#   HFTV1_PP_NCBI_overlap_df,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
#     "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
#     "b_PP_NCBI_OVERLAPS/HFTV1_PP_NCBI_overlap.csv", sep = ""))

# NCBI as query
HFTV1_NCBI_PP_overlap_df <- 
  inner_join(
    HFTV1_NCBI_CDS_df, 
    HFTV1_NCBI_PP_hits_df[,c("start", "end")], 
    by = c("start", "end"))

HFTV1_NCBI_PP_overlap_line <-
  paste(
    "Number of distinct overlapping features (NCBI as query): ",
    n_distinct(HFTV1_NCBI_PP_overlap_df$gene_id), sep = "")

HFTV1_NCBI_PP_overlap_gr <- 
  makeGRangesFromDataFrame(HFTV1_NCBI_PP_overlap_df, keep.extra.columns = TRUE)

# export(
#   HFTV1_NCBI_PP_overlap_gr,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
#     "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
#     "b_PP_NCBI_OVERLAPS/HFTV1_NCBI_PP_overlap.gff", sep = ""),
#   "gff3")
# write_csv(
#   HFTV1_NCBI_PP_overlap_df,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
#     "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
#     "b_PP_NCBI_OVERLAPS/HFTV1_NCBI_PP_overlap.csv", sep = ""))

```

# === HFTV1 | UNIQUE FEATURES OF NCBI AND P/P ================================
```{r}

# === Compile features unique to P/P prediction ===

HFTV1_PP_unique_df <-
  anti_join(
    HFTV1_PP_gff_df,
    HFTV1_PP_NCBI_hits_df[,c("start", "end")],
    by = c("start", "end"))

# Reformat unique P/P features to meet GTF column formatting requirements
HFTV1_PP_unique_df <-
  HFTV1_PP_unique_df %>%
  mutate(
    source = "Pharokka_Phold",
    # Adjust end position; standard GTF format requires CDS end to exclude
    # stop codon, whereas default Pharokka/Phold output includes stop codon.
    end = end - 3) %>%
  # Remove columns
  select(!width) %>%
  # Rename columns
  dplyr::rename(
    seqname = seqnames,
    feature = type,
    frame = phase) %>%
  # Relocate columns
  relocate(source, .after = seqname) %>%
  relocate(feature, .after = source) %>%
  relocate(start, .after = feature) %>%
  relocate(end, .after = start) %>%
  relocate(score, .after = end) %>%
  relocate(strand, .after = score) %>%
  relocate(frame, .after = strand)

HFTV1_PP_unique_line <-
  paste(
    "Number of distinct features unique to P/P: ",
    n_distinct(HFTV1_PP_unique_df$ID), sep = "")

HFTV1_PP_unique_gr <-
  makeGRangesFromDataFrame(HFTV1_PP_unique_df, keep.extra.columns = TRUE)

# export(
#   HFTV1_PP_unique_gr,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
#     "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
#     "c_PP_NCBI_UNIQUE/HFTV1_PP_unique.gff", sep = ""),
#   "gff3")
# write_csv(
#   HFTV1_PP_unique_df,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
#     "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
#     "c_PP_NCBI_UNIQUE/HFTV1_PP_unique.csv", sep = ""))

# === Compile features unique to NCBI ===

HFTV1_NCBI_unique_df <-
  anti_join(
    HFTV1_NCBI_CDS_df,
    HFTV1_NCBI_PP_hits_df[,c("start", "end")],
    by = c("start", "end"))

HFTV1_NCBI_unique_line <-
  paste(
    "Number of distinct features unique to NCBI: ",
    n_distinct(HFTV1_NCBI_unique_df$gene_id), sep = "")

HFTV1_NCBI_unique_gr <-
  makeGRangesFromDataFrame(HFTV1_NCBI_unique_df, keep.extra.columns = TRUE)

# export(
#   HFTV1_NCBI_unique_gr,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
#     "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
#     "c_PP_NCBI_UNIQUE/HFTV1_NCBI_unique.gff", sep = ""),
#   "gff3")
# write_csv(
#   HFTV1_NCBI_unique_df,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
#     "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
#     "c_PP_NCBI_UNIQUE/HFTV1_NCBI_unique.csv", sep = ""))


```

# === HFTV1 | NCBI-P/P REPORT ================================================
```{r}
# # Write various useful statistics to report file
# report_file <-
#   file(
#     paste(
#       "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
#       "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
#       "d_PP_NCBI_SUMMARY/report.txt", sep=""))
# # Write lines to file
# writeLines(
#   c(
#     HFTV1_NCBI_CDS_ct_line,
#     HFTV1_PP_ct_line,
#     HFTV1_PP_CDS_line,
#     HFTV1_PP_tRNA_line,
#     HFTV1_PP_NCBI_overlap_line,
#     HFTV1_NCBI_PP_overlap_line,
#     HFTV1_PP_unique_line,
#     HFTV1_NCBI_unique_line),
#   report_file)
# # Close file
# close(report_file)
```

# === HFTV1 | REFORMATTING UNIQUE P/P FEATURES FOR GTF =======================
```{r}
HFTV1_PP_unique_df <-
  HFTV1_PP_unique_df %>%
  mutate(
    # Rename tRNA feature type to "transcript" per Hg NCBI format. 
    feature = str_replace(
      feature, 
      pattern = "tRNA", 
      replacement = "transcript")) %>%
  # Rename columns to match NCBI format
  dplyr::rename(gene_id = ID) %>%
  dplyr::rename(transcript_id = Name) %>%
  # Remove certain columns to match NCBI format
  select(!c(
    function., phrog, source.1, translation, anticodon, isotype, trna)) %>%
  # Create new columns to match NCBI format
  mutate(
    gene_biotype = NA,
    transcript_id = paste("unassigned_transcript_", transcript_id, sep = ""),
    db_xref = paste("GeneID:MFI_PP_", gene_id, sep = ""),
    gbkey = NA,
    locus_tag = gene_id,
    old_locus_tag = NA,
    protein_id = NA,
    note = NA,
    exon_number = NA) %>%
  # Set frame to NA for tRNA
  mutate(frame = case_when(feature == "transcript" ~ NA, TRUE ~ frame)) %>%
  relocate(db_xref, .after = transcript_id) %>%
  relocate(gbkey, .after = db_xref) %>%
  relocate(gene_biotype, .after = gbkey) %>%
  relocate(locus_tag, .after = gene_biotype) %>%
  relocate(old_locus_tag, .after = locus_tag) %>%
  relocate(product, .after = old_locus_tag) %>%
  relocate(protein_id, .after = product) %>%
  relocate(transl_table, .after = protein_id) %>%
  relocate(exon_number, .after = transl_table) %>%
  relocate(note, .after = exon_number)

HFTV1_PP_unique_df$source <- as.factor(HFTV1_PP_unique_df$source)
  
# Subset unique P/P features to CDSs
HFTV1_PP_unique_CDS_df <-
  HFTV1_PP_unique_df %>%
  filter(feature == "CDS")

# === GTF formatting for P/P features

# Make gene entries for CDSs
HFTV1_PP_unique_gene_entries_df <-
  HFTV1_PP_unique_CDS_df %>%
  mutate(
    feature = "gene",
    frame = NA,
    end = end + 3,
    gene_biotype = "protein_coding",
    transcript_id = "")
# Make start codon entries
HFTV1_PP_unique_start_entries_df <-
  HFTV1_PP_unique_CDS_df %>%
  mutate(
    feature = "start_codon",
    end = start + 2)
# Make stop codon entries
HFTV1_PP_unique_stop_entries_df <-
  HFTV1_PP_unique_CDS_df %>%
  mutate(
    feature = "stop_codon",
    start = end + 1,
    end = end + 3)

# Make gene entries for tRNA
HFTV1_PP_unique_tRNA_gene_entries_df <-
  HFTV1_PP_unique_df %>% 
  filter(feature == "transcript") %>%
  mutate(
    feature = "gene",
    frame = NA)
# Make exon entries for tRNA
HFTV1_PP_unique_tRNA_exon_entries_df <-
  HFTV1_PP_unique_df %>% 
  filter(feature == "transcript") %>%
  mutate(
    feature = "exon",
    frame = NA)

# Combine all features together
HFTV1_PP_unique_gtf_df <-
  rbind(
    # All P/P unique features, including both CDSs and the tRNA
    HFTV1_PP_unique_df,
    # All gene entries for CDSs
    HFTV1_PP_unique_gene_entries_df,
    # All gene entries for tRNA (should be just one)
    HFTV1_PP_unique_tRNA_gene_entries_df,
    # All exon entries for tRNA (should be just one)
    HFTV1_PP_unique_tRNA_exon_entries_df,
    # All start codon entries for CDSs
    HFTV1_PP_unique_start_entries_df,
    # All stop codon entries for CDSs
    HFTV1_PP_unique_stop_entries_df)

# Set factor levels of features
HFTV1_PP_unique_gtf_df$feature <-
  factor(
    HFTV1_PP_unique_gtf_df$feature,
    levels = c(
      "gene", "CDS", "transcript", "exon", "start_codon", "stop_codon"))

# Revise columns to be consistent with GTF format
HFTV1_PP_unique_gtf_df %>%
  mutate(
    gbkey = case_when(
      feature == "gene" ~ "Gene",
      feature == "CDS" ~ "CDS",
      feature == "transcript" ~ "tRNA",
      feature == "exon" ~ NA,
      TRUE ~ NA))

# Arrange GTF by start position and feature (factor)
HFTV1_PP_unique_gtf_df <-
  HFTV1_PP_unique_gtf_df %>%
  arrange(start, feature)

HFTV1_PP_unique_gtf_gr <-
  makeGRangesFromDataFrame(
    HFTV1_PP_unique_gtf_df, keep.extra.columns = TRUE)

HFTV1_PP_unique_genes_gtf_gr <-
  makeGRangesFromDataFrame(
    HFTV1_PP_unique_gtf_df %>% 
      filter(feature == "gene"), keep.extra.columns = TRUE)

# === EXPORT ===

export(
  HFTV1_PP_unique_gtf_gr,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
    "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
    "c_PP_NCBI_UNIQUE/HFTV1_PP_unique_reformat.gtf", sep = ""),
  "gff3")

export(
  HFTV1_PP_unique_genes_gtf_gr,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
    "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
    "c_PP_NCBI_UNIQUE/HFTV1_PP_unique_genes_reformat.gtf", sep = ""),
  "gff3")

write_csv(
  HFTV1_PP_unique_gtf_df,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
    "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
    "c_PP_NCBI_UNIQUE/HFTV1_PP_unique_reformat.csv", sep = ""))
```

# === HFTV1 | CONSENSUS GTF OF NCBI AND UNIQUE P/P FEATURES ==================
```{r}

# Combine NCBI and unique P/P features
HFTV1_NCBI_PP_gtf_df <-
  rbind(
    HFTV1_NCBI_gtf_df,
    HFTV1_PP_unique_gtf_df) %>%
  mutate(Name = gene_id)

# Arrange GTF by mean start position per locus tag, then by feature (factor);
# add and then remove mean start position column. 
HFTV1_NCBI_PP_gtf_df <-
  HFTV1_NCBI_PP_gtf_df %>%
  group_by(locus_tag) %>%
  dplyr::mutate(mean_start = mean(start)) %>%
  ungroup() %>%
  arrange(mean_start, locus_tag, feature) %>%
  select(!mean_start)

# Convert GTF to Granges object for export
HFTV1_NCBI_PP_gtf_gr <-
  makeGRangesFromDataFrame(
    HFTV1_NCBI_PP_gtf_df, keep.extra.columns = TRUE)

# Subset GTF to "gene" entries only (easier for visualization in IGV), then
# convert GTF to Granges object for export.
HFTV1_NCBI_PP_genes_gtf_gr <-
  makeGRangesFromDataFrame(
    HFTV1_NCBI_PP_gtf_df %>% 
      filter(feature == "gene"), keep.extra.columns = TRUE)

export(
  HFTV1_NCBI_PP_gtf_gr,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
    "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
    "e_PP_NCBI_CONSENSUS/HFTV1_NCBI_PP_consensus.gtf", sep = ""),
  "gff3")

write_csv(
  HFTV1_NCBI_PP_gtf_df,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
    "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
    "e_PP_NCBI_CONSENSUS/HFTV1_NCBI_PP_consensus.csv", sep = ""))

export(
  HFTV1_NCBI_PP_genes_gtf_gr,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
    "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
    "e_PP_NCBI_CONSENSUS/HFTV1_NCBI_PP_genes_consensus.gtf", sep = ""),
  "gff3")

write_csv(
  HFTV1_NCBI_PP_gtf_df %>%
    filter(feature == "gene"),
  paste(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
    "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
    "e_PP_NCBI_CONSENSUS/HFTV1_NCBI_PP_genes_consensus.csv", sep = ""))
```

# === HFTV1 | RENAMING BASED ON TQ RNA-SEQ PAPER =============================
```{r}
# Rename genes with more standardized format
HFTV1_NCBI_PP_newnames_gtf_df <-
  HFTV1_NCBI_PP_gtf_df %>%
  mutate(
    gene_id = case_when(
      gene_id == "MNCMNMYK_CDS_0006" ~ "M1M17_gp06a",
      gene_id == "MNCMNMYK_CDS_0014" ~ "M1M17_gp13a",
      gene_id == "MNCMNMYK_CDS_0027" ~ "M1M17_gp25a",
      gene_id == "MNCMNMYK_CDS_0036" ~ "M1M17_gp33a",
      gene_id == "MNCMNMYK_CDS_0044" ~ "M1M17_gp39a",
      gene_id == "MNCMNMYK_CDS_0050" ~ "M1M17_gp46c",
      gene_id == "MNCMNMYK_CDS_0051" ~ "M1M17_gp46a",
      gene_id == "MNCMNMYK_CDS_0052" ~ "M1M17_gp46b",
      gene_id == "MNCMNMYK_CDS_0058" ~ "M1M17_gp50b",
      gene_id == "MNCMNMYK_CDS_0059" ~ "M1M17_gp50c",
      gene_id == "MNCMNMYK_CDS_0060" ~ "M1M17_gp52a",
      gene_id == "MNCMNMYK_CDS_0061" ~ "M1M17_gp52b",
      gene_id == "MNCMNMYK_CDS_0071" ~ "M1M17_gp61a",
      gene_id == "MNCMNMYK_CDS_0075" ~ "M1M17_gp64a",
      gene_id == "MNCMNMYK_CDS_0078" ~ "M1M17_gp66a",
      gene_id == "MNCMNMYK_CDS_0080" ~ "M1M17_gp67a",
      gene_id == "MNCMNMYK_tRNA_0001" ~ "M1M17_tRNA01",
      TRUE ~ gene_id),
    Name = case_when(
      Name == "MNCMNMYK_CDS_0006" ~ "M1M17_gp06a",
      Name == "MNCMNMYK_CDS_0014" ~ "M1M17_gp13a",
      Name == "MNCMNMYK_CDS_0027" ~ "M1M17_gp25a",
      Name == "MNCMNMYK_CDS_0036" ~ "M1M17_gp33a",
      Name == "MNCMNMYK_CDS_0044" ~ "M1M17_gp39a",
      Name == "MNCMNMYK_CDS_0050" ~ "M1M17_gp46c",
      Name == "MNCMNMYK_CDS_0051" ~ "M1M17_gp46a",
      Name == "MNCMNMYK_CDS_0052" ~ "M1M17_gp46b",
      Name == "MNCMNMYK_CDS_0058" ~ "M1M17_gp50b",
      Name == "MNCMNMYK_CDS_0059" ~ "M1M17_gp50c",
      Name == "MNCMNMYK_CDS_0060" ~ "M1M17_gp52a",
      Name == "MNCMNMYK_CDS_0061" ~ "M1M17_gp52b",
      Name == "MNCMNMYK_CDS_0071" ~ "M1M17_gp61a",
      Name == "MNCMNMYK_CDS_0075" ~ "M1M17_gp64a",
      Name == "MNCMNMYK_CDS_0078" ~ "M1M17_gp66a",
      Name == "MNCMNMYK_CDS_0080" ~ "M1M17_gp67a",
      Name == "MNCMNMYK_tRNA_0001" ~ "M1M17_tRNA01",
      TRUE ~ Name),
    locus_tag = case_when(
      locus_tag == "MNCMNMYK_CDS_0006" ~ "M1M17_gp06a",
      locus_tag == "MNCMNMYK_CDS_0014" ~ "M1M17_gp13a",
      locus_tag == "MNCMNMYK_CDS_0027" ~ "M1M17_gp25a",
      locus_tag == "MNCMNMYK_CDS_0036" ~ "M1M17_gp33a",
      locus_tag == "MNCMNMYK_CDS_0044" ~ "M1M17_gp39a",
      locus_tag == "MNCMNMYK_CDS_0050" ~ "M1M17_gp46c",
      locus_tag == "MNCMNMYK_CDS_0051" ~ "M1M17_gp46a",
      locus_tag == "MNCMNMYK_CDS_0052" ~ "M1M17_gp46b",
      locus_tag == "MNCMNMYK_CDS_0058" ~ "M1M17_gp50b",
      locus_tag == "MNCMNMYK_CDS_0059" ~ "M1M17_gp50c",
      locus_tag == "MNCMNMYK_CDS_0060" ~ "M1M17_gp52a",
      locus_tag == "MNCMNMYK_CDS_0061" ~ "M1M17_gp52b",
      locus_tag == "MNCMNMYK_CDS_0071" ~ "M1M17_gp61a",
      locus_tag == "MNCMNMYK_CDS_0075" ~ "M1M17_gp64a",
      locus_tag == "MNCMNMYK_CDS_0078" ~ "M1M17_gp66a",
      locus_tag == "MNCMNMYK_CDS_0080" ~ "M1M17_gp67a",
      locus_tag == "MNCMNMYK_tRNA_0001" ~ "M1M17_tRNA01",
      TRUE ~ locus_tag)) %>%
  mutate(
    Parent = paste("gene-", gene_id, sep = ""),
    ID = paste(str_to_lower(feature), "-", Name, sep = ""))

# Convert GTF to Granges object for export
HFTV1_NCBI_PP_newnames_gtf_gr <-
  makeGRangesFromDataFrame(
    HFTV1_NCBI_PP_newnames_gtf_df, keep.extra.columns = TRUE)

# === EXPORT ===

export(
  HFTV1_NCBI_PP_newnames_gtf_gr,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
    "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
    "f_PP_NCBI_CONSENSUS_NEWNAMES/HFTV1_NCBI_PP_consensus.gtf",
    sep = ""),
  "gff3")
write_csv(
  HFTV1_NCBI_PP_newnames_gtf_df,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
    "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
    "f_PP_NCBI_CONSENSUS_NEWNAMES/HFTV1_NCBI_PP_consensus_gtf.csv",
    sep = ""))
```

# === HFTV1 | RENAMING BASED ON TQ RNA-SEQ PAPER | GTF TO GFF ================
```{r}
# Make adjustments for GFF, then convert to GRanges object
HFTV1_NCBI_PP_newnames_gff_df <-
  HFTV1_NCBI_PP_newnames_gtf_df %>%
  dplyr::rename(
    type = feature,
    phase = frame) %>%
  filter(type != "start_codon") %>%
  filter(type != "stop_codon") %>%
  mutate(
    width = abs(end - start) + 1,
    end = case_when(
      type == "CDS" ~ end + 3,
      TRUE ~ end)) %>%
  select(!gene_id) %>%
  select(!transcript_id)
HFTV1_NCBI_PP_newnames_gff_gr <-
  makeGRangesFromDataFrame(
    HFTV1_NCBI_PP_newnames_gff_df, keep.extra.columns = TRUE)

# === EXPORT ===

# GFF for easy visualization in IGV
export(
  HFTV1_NCBI_PP_newnames_gff_gr,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
    "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
    "f_PP_NCBI_CONSENSUS_NEWNAMES/HFTV1_NCBI_PP_consensus.gff",
    sep = ""),
  "gff3")
write_csv(
  HFTV1_NCBI_PP_newnames_gff_df,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
    "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
    "f_PP_NCBI_CONSENSUS_NEWNAMES/HFTV1_NCBI_PP_consensus_gff.csv",
    sep = ""))
write_csv(
  HFTV1_NCBI_PP_newnames_gff_df %>% filter(type == "CDS"),
  paste(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
    "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
    "f_PP_NCBI_CONSENSUS_NEWNAMES/HFTV1_NCBI_PP_consensus_gff_cds.csv",
    sep = ""))


```

# === NOTES ON INCOMPATABILITIES BETWEEN GTFs AND Rtracklayer ================
```{r}
# Notes on issues with handling GTFs through Rtracklayer:
# https://support.bioconductor.org/p/70054/
```

# === HFTV1 | +/- HHT | REVISED ANNOTATION | CONSTRUCT + EXPORT ==============
```{r}
# Working with GFF now because my exported GTFs from above are really in
# GFF format (Rtracklayer apparently squeezes everything into GFF format, 
# and GFF v2 seems to differ from GTF as far as I can tell, despite
# contradictory reports online).

# Remove gp46c, gp52a, and gp52b - these were P/P-based predicted features
# unique to my analysis, and they showed very little evidence of TSSs in my
# HHT data. 

# Other adjustments (N-terminal extensions, N-terminal truncations, and
# addition of novel ORF - are based on manual inspection of +/-HHT data 
# in IGV).

HFTV1_NCBI_PP_gff_gr <-
  rtracklayer::import(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
      "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
      "f_PP_NCBI_CONSENSUS_NEWNAMES/HFTV1_NCBI_PP_consensus.gff",
      sep = ""))
HFTV1_NCBI_PP_gff_df <- 
  as.data.frame(HFTV1_NCBI_PP_gff_gr) %>%
  rowwise() %>%
  dplyr::mutate(Parent = paste(Parent, collapse="")) %>%
  ungroup()

HFTV1_NCBI_PP_HHT_gff_df <-
  HFTV1_NCBI_PP_gff_df %>%
  mutate(
    start = case_when(
      # N-terminal extensions
      locus_tag == "M1M17_gp12" ~ start - 5*3,
      locus_tag == "M1M17_gp21" ~ start - 8*3,
      locus_tag == "M1M17_gp19" ~ start - 22*3,
      locus_tag == "M1M17_gp30" ~ start - 50*3,
      locus_tag == "M1M17_gp66a" ~ start - 10*3,
      # N-terminal truncations
      locus_tag == "M1M17_gp05" ~ start + 18*3,
      locus_tag == "M1M17_gp56" ~ start + 39*3,
      locus_tag == "M1M17_gp67" ~ start + 5*3,
      TRUE ~ start)) %>%
  
  # Add novel features
  
  # gp52c
  add_row(
    seqnames = "NC_062739.1",
    start = 28684,
    end = 29004,
    width = end - start + 1,
    strand = "+",
    source = "JDR_HHT",
    type = "gene",
    score = NA,
    phase = NA,
    db_xref = NA,
    gbkey = "Gene",
    gene_biotype = "protein_coding",
    locus_tag = "M1M17_gp52c",
    old_locus_tag = NA,
    Name = "M1M17_gp52c",
    Parent = "gene-M1M17_gp52c",
    ID = "gene-M1M17_gp52c",
    product = NA,
    protein_id = NA,
    transl_table = NA,
    exon_number = NA,
    note = NA) %>%
  add_row(
    seqnames = "NC_062739.1",
    start = 28684,
    end = 29004,
    width = end - start + 1,
    strand = "+",
    source = "JDR_HHT",
    type = "CDS",
    score = NA,
    phase = 0,
    db_xref = NA,
    gbkey = "CDS",
    gene_biotype = "NA",
    locus_tag = "M1M17_gp52c",
    old_locus_tag = NA,
    Name = "M1M17_gp52c",
    Parent = "gene-M1M17_gp52c",
    ID = "cds-M1M17_gp52c",
    # Product based on Interpro-predicted domains
    product = "winged helix protein",
    protein_id = NA,
    transl_table = NA,
    exon_number = NA,
    note = NA) %>%
  
  # Remove P/P features that were not detected based on normalized HHT data
  
  filter(!Name == "M1M17_gp46c") %>%
  filter(!Name == "M1M17_gp52a") %>%
  filter(!Name == "M1M17_gp52b")
  
HFTV1_NCBI_PP_HHT_gff_df <-
  HFTV1_NCBI_PP_HHT_gff_df %>%
  group_by(locus_tag) %>%
  dplyr::mutate(mean_start = mean(start)) %>%
  ungroup() %>%
  arrange(mean_start, locus_tag, type) %>%
  select(!mean_start)

HFTV1_NCBI_PP_HHT_gff_gr <-
  makeGRangesFromDataFrame(
    HFTV1_NCBI_PP_HHT_gff_df, keep.extra.columns = TRUE)

export(
  HFTV1_NCBI_PP_HHT_gff_gr,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/h_REVISED_ANNOTATIONS/",
    "HFTV1_NCBI_PP_HHT_consensus.gff",
    sep = ""),
  "gff3")

write_csv(
  HFTV1_NCBI_PP_HHT_gff_df,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/h_REVISED_ANNOTATIONS/",
    "HFTV1_NCBI_PP_HHT_consensus_gff.csv",
    sep = ""))
```

# === HFTV1 | +/- HHT | REVISED ANNOTATION | HG MERGE + EXPORT ===============
```{r}
# Convert Hg NCBI annotation to the same GFF format as splitgp17
# Make adjustments for GFF, then convert to GRanges object
Hg_NCBI_gff_df <-
  Hg_NCBI_gtf_df %>%
  dplyr::rename(
    type = feature,
    phase = frame,
    seqnames = seqname) %>%
  filter(type != "start_codon") %>%
  filter(type != "stop_codon") %>%
  mutate(Name = locus_tag) %>%
  mutate(
    width = abs(end - start) + 1,
    end = case_when(
      type == "CDS" ~ end + 3,
      TRUE ~ end)) %>%
  mutate(
    Parent = paste("gene-", gene_id, sep = ""),
    ID = paste(str_to_lower(type), "-", Name, sep = "")) %>%
  select(c(
    seqnames, start, end, width, strand, source, type, score, phase,
    db_xref, gbkey, gene_biotype, locus_tag, old_locus_tag, Name, Parent,
    ID, product, protein_id, transl_table, exon_number, note))

HFTV1_Hg_gff_df <-
  rbind(
    HFTV1_NCBI_PP_HHT_gff_df,
    Hg_NCBI_gff_df)

HFTV1_Hg_gff_gr <-
  makeGRangesFromDataFrame(
    HFTV1_Hg_gff_df, keep.extra.columns = TRUE)

export(
  HFTV1_Hg_gff_gr,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/h_REVISED_ANNOTATIONS/",
    "HFTV1_Hg.gff",
    sep = ""),
  "gff3")

write_csv(
  HFTV1_Hg_gff_df,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/h_REVISED_ANNOTATIONS/",
    "HFTV1_Hg_gff.csv",
    sep = ""))
```

# === HFTV1 | +/- HHT | REVISED ANNOTATION | CONSTRUCT + EXPORT AND HG MERGE
# + EXPORT | INCLUDE LOW-CONFIDENCE FEATURES FOR MS ==========================
```{r}
# Same as above, but retain low-conf features for search in MS data. 

HFTV1_NCBI_PP_gff_gr <-
  rtracklayer::import(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
      "GCF_004208775.1/ANNOT_FUNCTION_REVISION/PHOLD_ANNOT/",
      "f_PP_NCBI_CONSENSUS_NEWNAMES/HFTV1_NCBI_PP_consensus.gff",
      sep = ""))

HFTV1_NCBI_PP_gff_df <- 
  as.data.frame(HFTV1_NCBI_PP_gff_gr) %>%
  rowwise() %>%
  dplyr::mutate(Parent = paste(Parent, collapse="")) %>%
  ungroup()

HFTV1_NCBI_PP_HHT_gff_df <-
  HFTV1_NCBI_PP_gff_df %>%
  mutate(
    start = case_when(
      # N-terminal extensions
      locus_tag == "M1M17_gp12" ~ start - 5*3,
      locus_tag == "M1M17_gp21" ~ start - 8*3,
      locus_tag == "M1M17_gp19" ~ start - 22*3,
      locus_tag == "M1M17_gp30" ~ start - 50*3,
      locus_tag == "M1M17_gp66a" ~ start - 10*3,
      # N-terminal truncations
      locus_tag == "M1M17_gp05" ~ start + 18*3,
      locus_tag == "M1M17_gp56" ~ start + 39*3,
      locus_tag == "M1M17_gp67" ~ start + 5*3,
      TRUE ~ start)) %>%
  
  # Add novel features
  
  # gp52c
  add_row(
    seqnames = "NC_062739.1",
    start = 28684,
    end = 29004,
    width = end - start + 1,
    strand = "+",
    source = "JDR_HHT",
    type = "gene",
    score = NA,
    phase = NA,
    db_xref = NA,
    gbkey = "Gene",
    gene_biotype = "protein_coding",
    locus_tag = "M1M17_gp52c",
    old_locus_tag = NA,
    Name = "M1M17_gp52c",
    Parent = "gene-M1M17_gp52c",
    ID = "gene-M1M17_gp52c",
    product = NA,
    protein_id = NA,
    transl_table = NA,
    exon_number = NA,
    note = NA) %>%
  add_row(
    seqnames = "NC_062739.1",
    start = 28684,
    end = 29004,
    width = end - start + 1,
    strand = "+",
    source = "JDR_HHT",
    type = "CDS",
    score = NA,
    phase = 0,
    db_xref = NA,
    gbkey = "CDS",
    gene_biotype = "NA",
    locus_tag = "M1M17_gp52c",
    old_locus_tag = NA,
    Name = "M1M17_gp52c",
    Parent = "gene-M1M17_gp52c",
    ID = "cds-M1M17_gp52c",
    # Product based on Interpro-predicted domains
    product = "winged helix protein",
    protein_id = NA,
    transl_table = NA,
    exon_number = NA,
    note = NA)
  
HFTV1_NCBI_PP_HHT_gff_df <-
  HFTV1_NCBI_PP_HHT_gff_df %>%
  group_by(locus_tag) %>%
  dplyr::mutate(mean_start = mean(start)) %>%
  ungroup() %>%
  arrange(mean_start, locus_tag, type) %>%
  select(!mean_start)

HFTV1_NCBI_PP_HHT_gff_gr <-
  makeGRangesFromDataFrame(
    HFTV1_NCBI_PP_HHT_gff_df, keep.extra.columns = TRUE)

export(
  HFTV1_NCBI_PP_HHT_gff_gr,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/h_REVISED_ANNOTATIONS/MS_annots/",
    "HFTV1_NCBI_PP_HHT_consensus_forMS.gff",
    sep = ""),
  "gff3")

write_csv(
  HFTV1_NCBI_PP_HHT_gff_df,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/h_REVISED_ANNOTATIONS/MS_annots/",
    "HFTV1_NCBI_PP_HHT_consensus_gff_forMS.csv",
    sep = ""))

# MERGE WITH HG
# Convert Hg NCBI annotation to the same GFF format as splitgp17
# Make adjustments for GFF, then convert to GRanges object
Hg_NCBI_gff_df <-
  Hg_NCBI_gtf_df %>%
  dplyr::rename(
    type = feature,
    phase = frame,
    seqnames = seqname) %>%
  filter(type != "start_codon") %>%
  filter(type != "stop_codon") %>%
  mutate(Name = locus_tag) %>%
  mutate(
    width = abs(end - start) + 1,
    end = case_when(
      type == "CDS" ~ end + 3,
      TRUE ~ end)) %>%
  mutate(
    Parent = paste("gene-", gene_id, sep = ""),
    ID = paste(str_to_lower(type), "-", Name, sep = "")) %>%
  select(c(
    seqnames, start, end, width, strand, source, type, score, phase,
    db_xref, gbkey, gene_biotype, locus_tag, old_locus_tag, Name, Parent,
    ID, product, protein_id, transl_table, exon_number, note))

HFTV1_Hg_gff_df <-
  rbind(
    HFTV1_NCBI_PP_HHT_gff_df,
    Hg_NCBI_gff_df) 

HFTV1_Hg_gff_df$type <- 
  factor(
    HFTV1_Hg_gff_df$type, 
    levels = c("gene", "transcript", "CDS", "exon"))

HFTV1_Hg_gff_df <-
  HFTV1_Hg_gff_df %>%
  group_by(locus_tag) %>%
  # Calculate mean start position for arranging features
  dplyr::mutate(mean_start = mean(start)) %>%
  # Fill in db_xref values within each locus tag
  fill(db_xref, .direction = "downup") %>%
  ungroup() %>%
  arrange(seqnames, mean_start, locus_tag, type) %>%
  select(!mean_start)

HFTV1_Hg_gff_gr <-
  makeGRangesFromDataFrame(
    HFTV1_Hg_gff_df, keep.extra.columns = TRUE)

export(
  HFTV1_Hg_gff_gr,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/h_REVISED_ANNOTATIONS/MS_annots/",
    "HFTV1_Hg_forMS.gff",
    sep = ""),
  "gff3")

write_csv(
  HFTV1_Hg_gff_df,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/h_REVISED_ANNOTATIONS/MS_annots/",
    "HFTV1_Hg_gff_forMS.csv",
    sep = ""))
```

# === HFTV1 | +/- HHT | REVISED ANNOTATION | gp17 | SPLIT + EXPORT ===========
```{r}
# Create version of annotation with split gp17 feature
HFTV1_NCBI_PP_HHT_splitgp17_gff_df <-
  HFTV1_NCBI_PP_HHT_gff_df %>%
  
  # Remove gp17 entries
  filter(locus_tag != "M1M17_gp17") %>%
  
  # gp17_NTERM
  add_row(
    seqnames = "NC_062739.1",
    start = 7106,
    end = 7981,
    width = end - start + 1,
    strand = "+",
    source = "JDR_HHT",
    type = "gene",
    score = NA,
    phase = NA,
    db_xref = NA,
    gbkey = "Gene",
    gene_biotype = "protein_coding",
    locus_tag = "M1M17_gp17_NTERM",
    old_locus_tag = NA,
    Name = "M1M17_gp17_NTERM",
    Parent = "gene-M1M17_gp17_NTERM",
    ID = "gene-M1M17_gp17_NTERM",
    product = NA,
    protein_id = NA,
    transl_table = NA,
    exon_number = NA,
    note = NA) %>%
  add_row(
    seqnames = "NC_062739.1",
    start = 7106,
    end = 7981,
    width = end - start + 1,
    strand = "+",
    source = "JDR_HHT",
    type = "CDS",
    score = NA,
    phase = 0,
    db_xref = NA,
    gbkey = "CDS",
    gene_biotype = "NA",
    locus_tag = "M1M17_gp17_NTERM",
    old_locus_tag = NA,
    Name = "M1M17_gp17_NTERM",
    Parent = "gene-M1M17_gp17_NTERM",
    ID = "cds-M1M17_gp17_NTERM",
    product = "P2 gpO-like scaffolding/protease protein",
    protein_id = NA,
    transl_table = NA,
    exon_number = NA,
    note = NA) %>%


  # gp17_CTERM
  add_row(
    seqnames = "NC_062739.1",
    start = 7982,
    end = 8548,
    width = end - start + 1,
    strand = "+",
    source = "JDR_HHT",
    type = "gene",
    score = NA,
    phase = NA,
    db_xref = NA,
    gbkey = "Gene",
    gene_biotype = "protein_coding",
    locus_tag = "M1M17_gp17_CTERM",
    old_locus_tag = NA,
    Name = "M1M17_gp17_CTERM",
    Parent = "gene-M1M17_gp17_CTERM",
    ID = "gene-M1M17_gp17_CTERM",
    product = NA,
    protein_id = NA,
    transl_table = NA,
    exon_number = NA,
    note = NA) %>%
  add_row(
    seqnames = "NC_062739.1",
    start = 7982,
    end = 8548,
    width = end - start + 1,
    strand = "+",
    source = "JDR_HHT",
    type = "CDS",
    score = NA,
    phase = 0,
    db_xref = NA,
    gbkey = "CDS",
    gene_biotype = "NA",
    locus_tag = "M1M17_gp17_CTERM",
    old_locus_tag = NA,
    Name = "M1M17_gp17_CTERM",
    Parent = "gene-M1M17_gp17_CTERM",
    ID = "cds-M1M17_gp17_CTERM",
    product = "P2 gpO-like scaffolding/protease protein",
    protein_id = NA,
    transl_table = NA,
    exon_number = NA,
    note = NA)
  
# Arrange based on mean position of each feature
HFTV1_NCBI_PP_HHT_splitgp17_gff_df <-
  HFTV1_NCBI_PP_HHT_splitgp17_gff_df %>%
  group_by(locus_tag) %>%
  dplyr::mutate(mean_start = mean(start)) %>%
  ungroup() %>%
  arrange(mean_start, locus_tag, type) %>%
  select(!mean_start)
  # dplyr::rename(
    # feature = type) %>%
  # mutate(gene_id = locus_tag) %>%
  # relocate(gene_id, .after = phase)

HFTV1_NCBI_PP_HHT_splitgp17_gff_gr <-
  makeGRangesFromDataFrame(
    HFTV1_NCBI_PP_HHT_splitgp17_gff_df, keep.extra.columns = TRUE)

# === EXPORT ===

export(
  HFTV1_NCBI_PP_HHT_splitgp17_gff_gr,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17/c_ANNOT/",
    "HFTV1_NCBI_PP_HHT_splitgp17.gff",
    sep = ""),
  "gff3")

write_csv(
  HFTV1_NCBI_PP_HHT_splitgp17_gff_df,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17/c_ANNOT/",
    "HFTV1_NCBI_PP_HHT_splitgp17_gff.csv",
    sep = ""))
```

# === HFTV1 | +/- HHT | REVISED ANNOTATION | gp17 | HG MERGE + EXPORT ========
```{r}
HFTV1_splitgp17_Hg_gff_df <-
  rbind(
    HFTV1_NCBI_PP_HHT_splitgp17_gff_df,
    Hg_NCBI_gff_df)


HFTV1_NCBI_PP_HHT_splitgp17_gff_gr <-
  makeGRangesFromDataFrame(
    HFTV1_splitgp17_Hg_gff_df, keep.extra.columns = TRUE)

# === EXPORT ===

export(
  HFTV1_NCBI_PP_HHT_splitgp17_gff_gr,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17/c_ANNOT/",
    "HFTV1_gp17split_Hg.gff",
    sep = ""),
  "gff3")

write_csv(
  HFTV1_splitgp17_Hg_gff_df,
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17/c_ANNOT/",
    "HFTV1_gp17split_Hg_gff.csv",
    sep = ""))
```



