---
title: "HFTV1_gp17"
author: "mfisada"
date: "2026-07-22"
output: html_document
---

# === REMOVE OBJECTS (*) =====================================================
```{r}
rm(list = ls())
```

# === PACKAGES (*) ===========================================================
```{r}
library(rtracklayer)
library(tidyverse)
library(tidyr)
library(ivs) # For overlaps
library(data.table) # For conversion from df to df
library(Biostrings)
library(phylotools) # For reading FASTA files
library(seqinr) # For exporting FASTA files
library(plyr)
library(dplyr)
library(plyranges) # For reading bigwigs
```

# === HFTV1 | gp17 | FC FUN ==================================================
```{r}
fc_fn <- function(sample) {
  
  dataframe <- 
    # Read in featurecounts data
    read_tsv(
      paste(
        "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
        "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/",
        "d_FEATURECOUNTS/",
        sample, "_fc.txt", sep=""),
          # Skip first line (record of featurecounts command)
          skip=1, 
          # Do not show column type
          show_col_types = FALSE) %>%
    # Rename columns 
    dplyr::rename(
      # Column 1: gene ID
      gene_id = 1,
      # Column 7: sample-specific feature counts
      counts = 7) %>%
    # Select only the gene_id and read counts columns for further use. 
    # dplyr::select(c(1,7)) %>% 
    # Round read count values to nearest integer (for featurecount data
    # generated using --fraction option for reads that map to overlapping
    # features. 
    dplyr::mutate(across(where(is.numeric), round, 0)) %>%
    # Subset to unique entries
    unique()
  
  return(dataframe)
  
}

# Recall that BR2 RNA 180 and 060 were swapped
BR2_RNA_180_E_fc_splitgp17_df <- fc_fn("BR2_RNA_060_E")
BR3_RNA_180_E_fc_splitgp17_df <- fc_fn("BR3_RNA_180_E")
BR4_RNA_180_E_fc_splitgp17_df <- fc_fn("BR4_RNA_180_E")

BR2_RBO_180_E_R1_fc_splitgp17_df <- fc_fn("BR2_RBO_180_E_R1")
BR3_RBO_180_E_R1_fc_splitgp17_df <- fc_fn("BR3_RBO_180_E_R1")
BR4_RBO_180_E_R1_fc_splitgp17_df <- fc_fn("BR4_RBO_180_E_R1")
```

# === HFTV1 | gp17 | RNA EXPRESSION QUANTIFY =================================
```{r}
BR2_RNA_180_E_bpm_gr <-
  read_bigwig(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
      "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/e_COVERAGE/",
      "BR2_RNA_180_E_bpm.bw", sep = ""))
BR3_RNA_180_E_bpm_gr <-
  read_bigwig(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
      "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/e_COVERAGE/",
      "BR3_RNA_180_E_bpm.bw", sep = ""))
BR4_RNA_180_E_bpm_gr <-
  read_bigwig(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
      "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/e_COVERAGE/",
      "BR4_RNA_180_E_bpm.bw", sep = ""))

BR2_RNA_180_E_bpm_df <- as.data.frame(BR2_RNA_180_E_bpm_gr)
BR3_RNA_180_E_bpm_df <- as.data.frame(BR3_RNA_180_E_bpm_gr)
BR4_RNA_180_E_bpm_df <- as.data.frame(BR4_RNA_180_E_bpm_gr)


BR2_RNA_180_E_bpm_gp17_df <-
  BR2_RNA_180_E_bpm_df %>%
  filter(seqnames == "NC_062739.1") %>%
  filter(start > 7105) %>%
  filter(start < 8549) %>%
  mutate(
    region = case_when(
      start < 7982 ~ "N_TERM", 
      start > 7981 ~ "C_TERM", 
      TRUE ~ "NA")) %>%
  group_by(region) %>%
  dplyr::summarize(region_cumulative_bpms = sum(score)) %>%
  ungroup() %>%
  mutate(
    BR = "BR2",
    sample_type = "RNA-Seq")
BR3_RNA_180_E_bpm_gp17_df <-
  BR3_RNA_180_E_bpm_df %>%
  filter(seqnames == "NC_062739.1") %>%
  filter(start > 7105) %>%
  filter(start < 8549) %>%
  mutate(
    region = case_when(
      start < 7982 ~ "N_TERM", 
      start > 7981 ~ "C_TERM", 
      TRUE ~ "NA")) %>%
  group_by(region) %>%
  dplyr::summarize(region_cumulative_bpms = sum(score)) %>%
  ungroup() %>%
  mutate(
    BR = "BR3",
    sample_type = "RNA-Seq")
BR4_RNA_180_E_bpm_gp17_df <-
  BR4_RNA_180_E_bpm_df %>%
  filter(seqnames == "NC_062739.1") %>%
  filter(start > 7105) %>%
  filter(start < 8549) %>%
  mutate(
    region = case_when(
      start < 7982 ~ "N_TERM", 
      start > 7981 ~ "C_TERM", 
      TRUE ~ "NA")) %>%
  group_by(region) %>%
  dplyr::summarize(region_cumulative_bpms = sum(score)) %>%
  ungroup() %>%
  mutate(
    BR = "BR4",
    sample_type = "RNA-Seq")



```

# === HFTV1 | gp17 | RBO EXPRESSION QUANTIFY =================================
```{r}
BR2_RBO_180_E_R1_bpm_gr <-
  read_bigwig(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
      "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/e_COVERAGE/",
      "BR2_RBO_180_E_R1_bpm.bw", sep = ""))
BR3_RBO_180_E_R1_bpm_gr <-
  read_bigwig(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
      "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/e_COVERAGE/",
      "BR3_RBO_180_E_R1_bpm.bw", sep = ""))
BR4_RBO_180_E_R1_bpm_gr <-
  read_bigwig(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
      "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/e_COVERAGE/",
      "BR4_RBO_180_E_R1_bpm.bw", sep = ""))

BR2_RBO_180_E_R1_bpm_df <- as.data.frame(BR2_RBO_180_E_R1_bpm_gr)
BR3_RBO_180_E_R1_bpm_df <- as.data.frame(BR3_RBO_180_E_R1_bpm_gr)
BR4_RBO_180_E_R1_bpm_df <- as.data.frame(BR4_RBO_180_E_R1_bpm_gr)

BR2_RBO_180_E_R1_bpm_gp17_df <-
  BR2_RBO_180_E_R1_bpm_df %>%
  filter(seqnames == "NC_062739.1") %>%
  filter(start > 7105) %>%
  filter(start < 8549) %>%
  mutate(
    region = case_when(
      start < 7982 ~ "N_TERM", 
      start > 7981 ~ "C_TERM", 
      TRUE ~ "NA")) %>%
  group_by(region) %>%
  dplyr::summarize(region_cumulative_bpms = sum(score)) %>%
  ungroup() %>%
  mutate(
    BR = "BR2",
    sample_type = "Ribo-Seq")
BR3_RBO_180_E_R1_bpm_gp17_df <-
  BR3_RBO_180_E_R1_bpm_df %>%
  filter(seqnames == "NC_062739.1") %>%
  filter(start > 7105) %>%
  filter(start < 8549) %>%
  mutate(
    region = case_when(
      start < 7982 ~ "N_TERM", 
      start > 7981 ~ "C_TERM", 
      TRUE ~ "NA")) %>%
  group_by(region) %>%
  dplyr::summarize(region_cumulative_bpms = sum(score)) %>%
  ungroup() %>%
  mutate(
    BR = "BR3",
    sample_type = "Ribo-Seq")
BR4_RBO_180_E_R1_bpm_gp17_df <-
  BR4_RBO_180_E_R1_bpm_df %>%
  filter(seqnames == "NC_062739.1") %>%
  filter(start > 7105) %>%
  filter(start < 8549) %>%
  mutate(
    region = case_when(
      start < 7982 ~ "N_TERM", 
      start > 7981 ~ "C_TERM", 
      TRUE ~ "NA")) %>%
  group_by(region) %>%
  dplyr::summarize(region_cumulative_bpms = sum(score)) %>%
  ungroup() %>%
  mutate(
    BR = "BR4",
    sample_type = "Ribo-Seq")


```

# === HFTV1 | gp17 | COMBINE RNA AND RBO =====================================
```{r}

gp17_180_bpms_df <-
  rbind(
    BR2_RNA_180_E_bpm_gp17_df,
    BR3_RNA_180_E_bpm_gp17_df,
    BR4_RNA_180_E_bpm_gp17_df,
    BR2_RBO_180_E_R1_bpm_gp17_df,
    BR3_RBO_180_E_R1_bpm_gp17_df,
    BR4_RBO_180_E_R1_bpm_gp17_df)

gp17_180_bpms_df$sample_type <- 
  factor(
    gp17_180_bpms_df$sample_type,
    levels = c("RNA-Seq", "Ribo-Seq"))

gp17_180_bpms_df$region <- 
  factor(
    gp17_180_bpms_df$region,
    levels = c("N_TERM", "C_TERM"))

# write_csv(
#   gp17_180_bpms_df,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
#     "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17/e_COVERAGE/",
#     "gp17_180_bpms.csv", sep = ""))
```

# === HFTV1 | gp17 | RNA EXPRESSION QUANTIFY | VIS ===========================
```{r}

gp17_180_bpms_plot <-
  ggplot(
    gp17_180_bpms_df %>% filter(sample_type == "RNA-Seq"),
    aes(
      y = region_cumulative_bpms,
      x = sample_type,
      group = region,
      fill = region)) +
  
  geom_bar(
    stat = "summary",
    fun = "mean",
    position = "dodge",
    color="black", 
    size=0.75, 
    width=0.9) +
  
  geom_point(
    data = 
      gp17_180_bpms_df %>% 
      filter(sample_type == "RNA-Seq") %>%
      filter(BR == "BR2"),
    position = position_dodge(width = 0.9), 
    shape = 18, 
    size=3, 
    stroke=1.5) +
  
  geom_point(
    data = 
      gp17_180_bpms_df %>% 
      filter(sample_type == "RNA-Seq") %>%
      filter(BR == "BR3"),
    position = position_dodge(width = 0.9), 
    shape = 19, 
    size=1, 
    stroke=1.5) +
  
  geom_point(
    data = 
      gp17_180_bpms_df %>% 
      filter(sample_type == "RNA-Seq") %>%
      filter(BR == "BR4"),
    position = position_dodge(width = 0.9), 
    shape = 15, 
    size=2, 
    stroke=1.5) +
  
  stat_summary(
    fun.data = mean_sdl, 
    geom = "errorbar", 
    width=0.2, 
    position = position_dodge(width=0.9), fun.args = list(mult = 1)) +
  
  scale_fill_manual(
    values = c(
      C_TERM = "indianred3",
      N_TERM = "steelblue2")) +
  
  scale_y_continuous(
    labels = function(x) format(x, scientific = TRUE),
    limits = c(0, 1500000)) +
  
  ylab("Cumulative RPMs") +
  xlab("") +
  
  theme_classic() +
  theme(
    plot.title=element_text(size=18, face="bold", hjust = 0.5),

    axis.title = element_text(size = 16, face="bold"),
    axis.title.x = element_text(margin=margin(t=10)),
    axis.title.y = element_text(margin=margin(r=10)),

    axis.text = element_text(size = 14, face="bold", color="black"),

    axis.ticks = element_line(linewidth = 0.75, color = "black"),
    axis.ticks.length=unit(0.25, "cm"),

    axis.line=element_line(linewidth = 0, color = "black"),

    legend.text = element_text(face="bold", size=12),
    # No legend
    # legend.position="none",

    panel.background = element_rect(fill = "white"),
    panel.border =
      element_rect(colour = "black", fill=NA, linewidth=1.75),
    panel.spacing = unit(3, "lines"),

    strip.text.x =
      element_text(margin = margin(2, 0, 2, 0), size = 14, face="bold"),
    strip.background = element_blank())

# Export
ggsave(
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/e_COVERAGE/",
    "gp17_180_bpms_plot_RNA.png", sep = ""),
  plot = gp17_180_bpms_plot,
  width = 10,
  height = 8,
  units = "cm", 
  dpi = 300)

ggsave(
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/e_COVERAGE/",
    "gp17_180_bpms_plot_RNA.svg", sep = ""),
  plot = gp17_180_bpms_plot,
  width = 10,
  height = 8,
  units = "cm", 
  dpi = 300)
```

# === HFTV1 | gp17 | RBO EXPRESSION QUANTIFY | VIS ===========================
```{r}

gp17_180_bpms_plot <-
  ggplot(
    gp17_180_bpms_df %>% filter(sample_type == "Ribo-Seq"),
    aes(
      y = region_cumulative_bpms,
      x = sample_type,
      group = region,
      fill = region)) +
  
  geom_bar(
    stat = "summary",
    fun = "mean",
    position = "dodge",
    color="black", 
    size=0.75, 
    width=0.9) +
  
  geom_point(
    data = 
      gp17_180_bpms_df %>% 
      filter(sample_type == "Ribo-Seq") %>%
      filter(BR == "BR2"),
    position = position_dodge(width = 0.9), 
    shape = 18, 
    size=3, 
    stroke=1.5) +
  
  geom_point(
    data = 
      gp17_180_bpms_df %>% 
      filter(sample_type == "Ribo-Seq") %>%
      filter(BR == "BR3"),
    position = position_dodge(width = 0.9), 
    shape = 19, 
    size=1, 
    stroke=1.5) +
  
  geom_point(
    data = 
      gp17_180_bpms_df %>% 
      filter(sample_type == "Ribo-Seq") %>%
      filter(BR == "BR4"),
    position = position_dodge(width = 0.9), 
    shape = 15, 
    size=2, 
    stroke=1.5) +
  
  stat_summary(
    fun.data = mean_sdl, 
    geom = "errorbar", 
    width=0.2, 
    position = position_dodge(width=0.9), fun.args = list(mult = 1)) +
  
  scale_fill_manual(
    values = c(
      C_TERM = "indianred3",
      N_TERM = "steelblue2")) +
  
  scale_y_continuous(
    labels = function(x) format(x, scientific = TRUE),
    limits = c(0, 1500000)) +
  
  ylab("Cumulative RPMs") +
  xlab("") +
  
  theme_classic() +
  theme(
    plot.title=element_text(size=18, face="bold", hjust = 0.5),

    axis.title = element_text(size = 16, face="bold"),
    axis.title.x = element_text(margin=margin(t=10)),
    axis.title.y = element_text(margin=margin(r=10)),

    axis.text = element_text(size = 14, face="bold", color="black"),

    axis.ticks = element_line(linewidth = 0.75, color = "black"),
    axis.ticks.length=unit(0.25, "cm"),

    axis.line=element_line(linewidth = 0, color = "black"),

    legend.text = element_text(face="bold", size=12),
    # No legend
    # legend.position="none",

    panel.background = element_rect(fill = "white"),
    panel.border =
      element_rect(colour = "black", fill=NA, linewidth=1.75),
    panel.spacing = unit(3, "lines"),

    strip.text.x =
      element_text(margin = margin(2, 0, 2, 0), size = 14, face="bold"),
    strip.background = element_blank())

# Export
ggsave(
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/e_COVERAGE/",
    "gp17_180_bpms_plot_RBO.png", sep = ""),
  plot = gp17_180_bpms_plot,
  width = 10,
  height = 8,
  units = "cm", 
  dpi = 300)

ggsave(
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/e_COVERAGE/",
    "gp17_180_bpms_plot_RBO.svg", sep = ""),
  plot = gp17_180_bpms_plot,
  width = 10,
  height = 8,
  units = "cm", 
  dpi = 300)
```

# ratio wrangle
```{r}

gp17_180_CTERM_bpms_df <-
  gp17_180_bpms_df %>%
  filter(region == "C_TERM") %>%
  dplyr::rename(C_TERM_cum_bpms = region_cumulative_bpms)

gp17_180_NTERM_bpms_df <-
  gp17_180_bpms_df %>%
  filter(region == "N_TERM") %>%
  dplyr::rename(N_TERM_cum_bpms = region_cumulative_bpms)

gp17_180_ratio_df <-
  left_join(
    gp17_180_NTERM_bpms_df %>%
      select(!region),
    gp17_180_CTERM_bpms_df %>%
      select(!region),
    by = c("BR", "sample_type")) %>%
  relocate(N_TERM_cum_bpms, .before = C_TERM_cum_bpms) %>%
  mutate(ratio = C_TERM_cum_bpms/N_TERM_cum_bpms)

```

# ratio vis
```{r}
my_plot <-
  
  ggplot(
    gp17_180_ratio_df,
    aes(
      x = sample_type,
      y = ratio),
    group = sample_type) +
  
  geom_bar(
    stat = "summary",
    fun = "mean",
    position = "dodge",
    color="black", 
    size=0.75,
    width=0.9,
    fill = "grey90") +

  geom_point(
    data = gp17_180_ratio_df %>%
      filter(BR == "BR2"),
    aes(
      y = ratio,
      group = sample_type),
    position = position_dodge(width = 0.9), shape = 1, size=1, stroke = 1.25) +
  geom_point(
    data = gp17_180_ratio_df %>%
      filter(BR == "BR3"),
    aes(
      y = ratio,
      group = sample_type),
    position = position_dodge(width = 0.9), shape = 1, size=1, stroke = 1.25) +
  geom_point(
    data = gp17_180_ratio_df %>%
      filter(BR == "BR4"),
    aes(
      y = ratio,
      group = sample_type),
    position = position_dodge(width = 0.9), shape = 1, size=1, stroke = 1.25) +

  stat_summary(
    fun.data = mean_sdl,
    geom = "errorbar",
    width=0.2,
    position = position_dodge(width=0.9), fun.args = list(mult = 1)) +
  
  ylim(0, 10) +
  
  ylab("C-term.:N-term. cumulative RPMs") +

  theme_classic() +
  theme(
    plot.title=element_text(size=18, face="bold", hjust = 0.5),

    axis.title = element_text(size = 16, face="bold"),
    axis.title.x = element_text(margin=margin(t=10)),
    axis.title.y = element_text(margin=margin(r=10)),

    axis.text = element_text(size = 14, face="bold", color="black"),

    axis.ticks = element_line(linewidth = 0.75, color = "black"),
    axis.ticks.length=unit(0.25, "cm"),

    axis.line=element_line(linewidth = 0, color = "black"),

    legend.text = element_text(face="bold", size=12),
    # No legend
    # legend.position="none",

    panel.background = element_rect(fill = "white"),
    panel.border =
      element_rect(colour = "black", fill=NA, linewidth=1.75),
    panel.spacing = unit(3, "lines"),

    strip.text.x =
      element_text(margin = margin(2, 0, 2, 0), size = 14, face="bold"),
    strip.background = element_blank(),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    panel.grid.major.y = element_line(color = "grey80"),
    # panel.grid.major.x = element_line(color = "grey80"),
    panel.grid.minor.y = element_line(color = "grey80")) +
    # panel.grid.minor.x = element_line(color = "grey80"))
  
  facet_wrap(
    ~ sample_type,
    scales = "free"
  )


ggsave(
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/RNA_RBO/",
    "c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/e_COVERAGE/",
    "CTERM_NTERM_ratio.png",
    sep = ""),
  plot = my_plot,
  width = 12,
  height = 12,
  units = "cm", 
  dpi = 300)

ggsave(
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/RNA_RBO/",
    "c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/e_COVERAGE/",
    "CTERM_NTERM_ratio.svg",
    sep = ""),
  plot = my_plot,
  width = 12,
  height = 12,
  units = "cm", 
  dpi = 300)
```

# === gp17 | TE | ANNOTATIONS ================================================
```{r}
# === MERGED Hg-HFTV1 ANNOTATION ===

# Load in merged Hg-HFTV1 annotation
Hg_HFTV1_annot_gr <- 
  rtracklayer::import(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
      "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/c_ANNOT/",
      "HFTV1_gp17split_Hg.gff", 
      sep = ""))

# Convert GRanges to data.frame
Hg_HFTV1_annot_df <-  
  as.data.frame(Hg_HFTV1_annot_gr)%>%
  mutate(gene_id = locus_tag) %>%
  relocate(gene_id, .after = "Name")

# === PANTHER ===

# Load Hg PANTHER annotations
Hg_panther_df <- 
  read.delim(
    text = gsub(
      "\\t$", "",
      readLines(
        paste(
          "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/",
          "H_GIBBONSII/DL_20240429/GCF_014969745.1/",
          "PANTHER_ANNOTATION_20250328/Hg_PANTHER.txt", sep = ""))))
colnames(Hg_panther_df) <- 
  c("protein_id", "PANTHER_accession", "PANTHER_family", 
    "PANTHER_HMM_evalue", "PANTHER_HMM_bitscore", "PANTHER_alignment_range")

# Load HFTV1 PANTHER annotations
HFTV1_panther_df <- 
  read.delim(
    text = gsub(
      "\\t$", "",
      readLines(
        paste(
          "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/",
          "HFTV1/DL_20240430/GCF_004208775.1/ANNOT_FUNCTION_REVISION/",
          "PANTHER_ANNOT_20250501/HFTV1_PANTHER.txt", sep = ""))))
colnames(HFTV1_panther_df) <- 
  c("protein_id", "PANTHER_accession", "PANTHER_family", "PANTHER_HMM_evalue",
    "PANTHER_HMM_bitscore", "PANTHER_alignment_range")

# Combine Hg and HFTV1 PANTHER annotations
panther_df <- rbind(Hg_panther_df, HFTV1_panther_df)

# Remove semi-redundant annotations (multiple annotations per certain genes)
panther_df <-
  panther_df %>%
  filter(
    !(
      (protein_id == "WP_115804446.1" & 
         PANTHER_accession == "PTHR39576:SF2") |
        (protein_id == "WP_004974336.1" & 
           PANTHER_accession == "PTHR34236:SF1") |
        (protein_id == "WP_004974668.1" &
           PANTHER_accession == "PTHR10536:SF0") |
        (protein_id == "WP_004974320.1" &
           PANTHER_accession == "PTHR46229:SF2") |
        (protein_id == "WP_058827918.1" &
           PANTHER_accession == "PTHR34192:SF10")))

# === eggNOG ===

# Import COG code
COG_code_df <-
  read_csv(
    "/Users/mi/DIRUGGIERO_LAB/REFERENCES/COG_CODE/cog_code.csv",
    show_col_types = FALSE)

# Convert to factors to facilitate graphing much later
COG_code_df$COG_category <-
  factor(COG_code_df$COG_category,
         levels = COG_code_df$COG_category)
COG_code_df$COG_desc_lvl1 <-
  factor(COG_code_df$COG_desc_lvl1,
         levels = c(
           "information_storage_and_processing",
           "cellular_processes_and_signaling",
           "metabolism",
           "poorly_characterized",
           "NAN",
           "no_hits"))
COG_code_df$COG_desc_lvl1_long <-
  factor(COG_code_df$COG_desc_lvl1_long,
         levels = c(
           "Information storage and processing",
           "Cellular processes and signaling",
           "Metabolism",
           "Poorly characterized",
           "NAN", 
           "No hits"))
COG_code_df$COG_desc_lvl2 <-
  factor(COG_code_df$COG_desc_lvl2,
         levels = COG_code_df$COG_desc_lvl2)
COG_code_df$COG_desc_lvl2_long <-
  factor(COG_code_df$COG_desc_lvl2_long,
         levels = COG_code_df$COG_desc_lvl2_long)
  
# Import eggNOG-MAPPER annotations
Hg_eggNOG_df <- 
  read.table(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/H_GIBBONSII/DL_20250521/",
      "GCF_014969745.1/EGGNOG_MAPPER_20250905/annotations.tsv", sep = ""),
  quote = "",
  sep = "\t") %>%
  dplyr::rename(
    protein_id = V1, 
    eggNOG_seed_ortholog = V2, 
    eggNOG_evalue = V3, 
    eggNOG_score = V4, 
    eggNOG_OGs = V5, 
    eggNOG_max_annot_lvl = V6, 
    COG_category = V7, 
    Description = V8, 
    eggNOG_gene_name = V9, 
    eggNOG_GOs = V10, 
    eggNOG_EC = V11, 
    KEGG_ko = V12,
    KEGG_Pathway = V13, 
    KEGG_Module = V14, 
    KEGG_Reaction = V15,
    KEGG_rclass = V16, 
    BRITE = V17, 
    KEGG_TC = V18, 
    CAZy = V19,
    BIGG_Reaction = V20, 
    PFAMs = V21) 

HFTV1_eggNOG_df <- 
  read.table(
    paste(
      "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HFTV1/DL_20240430/",
      "GCF_004208775.1/ANNOT_FUNCTION_REVISION/EGGNOG_MAPPER_20250910/",
      "annotations.tsv", sep = ""),
  quote = "",
  sep = "\t") %>%
  dplyr::rename(
    protein_id = V1, 
    eggNOG_seed_ortholog = V2, 
    eggNOG_evalue = V3, 
    eggNOG_score = V4, 
    eggNOG_OGs = V5, 
    eggNOG_max_annot_lvl = V6, 
    COG_category = V7, 
    Description = V8, 
    eggNOG_gene_name = V9, 
    eggNOG_GOs = V10, 
    eggNOG_EC = V11, 
    KEGG_ko = V12,
    KEGG_Pathway = V13, 
    KEGG_Module = V14, 
    KEGG_Reaction = V15,
    KEGG_rclass = V16, 
    BRITE = V17, 
    KEGG_TC = V18, 
    CAZy = V19,
    BIGG_Reaction = V20, 
    PFAMs = V21) 

# Combine Hg and HFTV1 eggNOG information
Hg_HFTV1_eggNOG_df <-
  rbind(Hg_eggNOG_df, HFTV1_eggNOG_df)

# Combine eggNOG information with decoded COG information
Hg_HFTV1_eggNOG_df <- 
  left_join(
    Hg_HFTV1_eggNOG_df,
    COG_code_df,
    by = "COG_category")

# === COMBINING ANNOTATIONS ===

# Combine PANTHER annotations with NCBI-P/P-RPR annotations
Hg_HFTV1_annot_df <-
  merge(
    Hg_HFTV1_annot_df, 
    panther_df, 
    by="protein_id", all = TRUE)

# Combine eggNOG annotations with NCBI-P/P-RPR-PANTHER annotations
Hg_HFTV1_annot_df <-
  merge(
    Hg_HFTV1_annot_df, 
    Hg_HFTV1_eggNOG_df, 
    by="protein_id", all = TRUE) %>%
  mutate(
    # Convert NAs to string so that they can be manipulated more easily in
    # graphs later. 
    COG_category = replace_na(COG_category, "-"),
    COG_desc_lvl1 = replace_na(COG_desc_lvl1, "no_hits"),
    COG_desc_lvl1_long = replace_na(COG_desc_lvl1_long, "No hits"),
    COG_desc_lvl2 = replace_na(COG_desc_lvl2, "no_hits"),
    COG_desc_lvl2_long = replace_na(COG_desc_lvl2_long, "No hits"),
    
    # Manually reassign COG categories for genes with more than one
    # COG category.
    COG_category = replace(COG_category, COG_category == "GT", "G"),
    COG_category = replace(COG_category, COG_category == "EGP", "E"),
    COG_category = replace(COG_category, COG_category == "PT", "P"),
    COG_category = replace(COG_category, COG_category == "NU", "N"),
    COG_category = replace(COG_category, COG_category == "NT", "N"),
    COG_category = replace(COG_category, COG_category == "BQ", "B"),
    COG_category = replace(COG_category, COG_category == "JM", "J"),
    COG_category = replace(COG_category, COG_category == "CP", "C"),
    COG_category = replace(COG_category, COG_category == "GH", "G"))

# Add entity column to easily distinguish between host and virus
Hg_HFTV1_annot_df <-
  Hg_HFTV1_annot_df %>%
  mutate(
    entity = 
      case_when(str_detect(gene_id, "M1M17") ~ "HFTV1", TRUE ~ "Hg")) 

# Convert some COG columns to factors for graphing much further downstream
Hg_HFTV1_annot_df$COG_category <-
  factor(Hg_HFTV1_annot_df$COG_category,
         levels = COG_code_df$COG_category)
Hg_HFTV1_annot_df$COG_desc_lvl1 <-
  factor(Hg_HFTV1_annot_df$COG_desc_lvl1,
         levels = c(
           "information_storage_and_processing",
           "cellular_processes_and_signaling",
           "metabolism",
           "poorly_characterized",
           "NAN",
           "no_hits"))
Hg_HFTV1_annot_df$COG_desc_lvl1_long <-
  factor(Hg_HFTV1_annot_df$COG_desc_lvl1_long,
         levels = c(
           "Information storage and processing",
           "Cellular processes and signaling",
           "Metabolism",
           "Poorly characterized",
           "NAN", 
           "No hits"))
Hg_HFTV1_annot_df$COG_desc_lvl2 <-
  factor(Hg_HFTV1_annot_df$COG_desc_lvl2,
         levels = COG_code_df$COG_desc_lvl2)
Hg_HFTV1_annot_df$COG_desc_lvl2_long <-
  factor(Hg_HFTV1_annot_df$COG_desc_lvl2_long,
         levels = COG_code_df$COG_desc_lvl2_long)


# Subset to most useful columns
Hg_HFTV1_annot_df <-
  Hg_HFTV1_annot_df %>%
  select(c(
    entity,
    seqnames,
    gene_id,
    protein_id,
    start,
    end,
    width,
    strand,
    type,
    source,
    product,
    COG_category,
    COG_desc_lvl1,
    COG_desc_lvl1_long,
    COG_desc_lvl2,
    COG_desc_lvl2_long,
    eggNOG_gene_name,
    Description))


# === SUBSETS ===

# Create lists of all genes and of entity-specific genes
Hg_HFTV1_annot_genes <- unique(Hg_HFTV1_annot_df[ ,"gene_id"])
HFTV1_annot_genes <- Hg_HFTV1_annot_genes[grep("M1M17", (Hg_HFTV1_annot_genes))]
Hg_annot_genes <- Hg_HFTV1_annot_genes[grep("HfgLR", (Hg_HFTV1_annot_genes))]

# Subset to all non-p3 genes
Hg_HFTV1_annot_Np3_df <- 
  Hg_HFTV1_annot_df %>%
  filter(!grepl("NZ_CP063208.1", seqnames))
# Subset to all p3 genes
Hg_annot_p3_df <- 
  Hg_HFTV1_annot_df %>%
  filter(grepl("NZ_CP063208.1", seqnames))

# Subset entries to "gene" (feature level)
Hg_HFTV1_gene_df <- Hg_HFTV1_annot_df %>%
  filter(grepl("gene", type))
# Subset entries to "CDS" (metafeature level)
Hg_HFTV1_cds_df <- Hg_HFTV1_annot_df %>%
  filter(grepl("CDS", type))
# Subset entries to "tRNA" (metafeature level)
Hg_HFTV1_tRNA_df <- Hg_HFTV1_annot_df %>%
  filter(grepl("tRNA", type))

# Combine CDS and tRNA dataframes (two different types of metafeatures)
Hg_HFTV1_cds_tRNA_df <- rbind(Hg_HFTV1_cds_df, Hg_HFTV1_tRNA_df)

Hg_HFTV1_metafeat_annot_df <- 
  Hg_HFTV1_annot_df %>%
  # Subset entries to all types except "gene" (several types of metafeatures)
  filter(!grepl("gene", type)) %>%
  # Subset to remove exons as well, which are redundant with types tRNA,
  # rRNA, and RNase P RNA. 
  filter(!grepl("exon", type))


# Exports

# write_csv(
#   Hg_HFTV1_annot_df,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HG-HFTV1/RNA_RBO_ANNOTS/",
#     "Hg_HFTV1_annot.csv", sep = ""))
# write_csv(
#   Hg_HFTV1_metafeat_annot_df,
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HG-HFTV1/RNA_RBO_ANNOTS/",
#     "Hg_HFTV1_metafeat.csv", sep = ""))
# write_csv(
#   Hg_HFTV1_annot_df %>% filter(type == "pseudogene"),
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/REFERENCES/SEQ/HG-HFTV1/RNA_RBO_ANNOTS/",
#     "Hg_HFTV1_pseudogenes.csv", sep = ""))

```

# === COG FACTOR CONVERSION FUN (*) ==========================================
```{r}
COG_factor_conversion_fun <-
  function(
    input_df,
    COG_code = COG_code_df) {
    
    input_df$COG_category <-
      factor(
        input_df$COG_category,
        levels = COG_code$COG_category)
    
    input_df$COG_desc_lvl1 <-
      factor(
        input_df$COG_desc_lvl1,
        levels = c(
          "information_storage_and_processing",
           "cellular_processes_and_signaling",
           "metabolism",
           "poorly_characterized",
           "NAN",
           "no_hits"))
    
    input_df$COG_desc_lvl1_long <-
      factor(
        input_df$COG_desc_lvl1_long,
        levels = c(
          "Information storage and processing",
           "Cellular processes and signaling",
           "Metabolism",
           "Poorly characterized",
           "NAN", 
           "No hits"))
    
    input_df$COG_desc_lvl2 <-
      factor(
        input_df$COG_desc_lvl2,
        levels = COG_code_df$COG_desc_lvl2)
    
    input_df$COG_desc_lvl2_long <-
      factor(
        input_df$COG_desc_lvl2_long,
        levels = COG_code_df$COG_desc_lvl2_long)
    
    return(input_df)
    
  }


```

# === DEA | ISO | INPUTS =====================================================
```{r}
# Prepare data and metadata for DE analysis
# Some code is derived from SRC

# === RNA ===

RNA_fc_fn <- function(rep, sample) {
  
  dataframe <- 
    # Read in featurecounts data
    read_tsv(paste(
      "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
      "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/d_FEATURECOUNTS/",
      sample, "_fc.txt", sep=""),
             # Skip first line (record of featurecounts command)
             skip=1, 
             # Do not show column type
             show_col_types = FALSE) %>%
    # Rename columns 
    dplyr::rename(
      # Column 1: gene ID
      gene_id = 1,
      # Column 7: sample-specific feature counts
      {{sample}}:=7) %>%
    # Select only the gene_id and read counts columns for further use. 
    dplyr::select(c(1,7)) %>% 
    # Round read count values to nearest integer (for featurecount data
    # generated using --fraction option for reads that map to overlapping
    # features. 
    dplyr::mutate(across(where(is.numeric), round, 0)) %>%
    # Subset to unique entries
    unique()
  
  return(dataframe)
  
}

# BR2 (NOTE: account for sample swap, 180E <-> 060E)
BR2_RNA_000_E_fc_df <- RNA_fc_fn("BR2", "BR2_RNA_000_E")
BR2_RNA_005_E_fc_df <- RNA_fc_fn("BR2", "BR2_RNA_005_E")
BR2_RNA_015_E_fc_df <- RNA_fc_fn("BR2", "BR2_RNA_015_E")
BR2_RNA_030_E_fc_df <- RNA_fc_fn("BR2", "BR2_RNA_030_E")
BR2_RNA_060_E_fc_df <- RNA_fc_fn("BR2", "BR2_RNA_180_E") %>%
  dplyr::rename(BR2_RNA_060_E = BR2_RNA_180_E)
BR2_RNA_180_E_fc_df <- RNA_fc_fn("BR2", "BR2_RNA_060_E") %>%
  dplyr::rename(BR2_RNA_180_E = BR2_RNA_060_E)

# BR3
BR3_RNA_000_E_fc_df <- RNA_fc_fn("BR3", "BR3_RNA_000_E")
BR3_RNA_005_E_fc_df <- RNA_fc_fn("BR3", "BR3_RNA_005_E")
BR3_RNA_015_E_fc_df <- RNA_fc_fn("BR3", "BR3_RNA_015_E")
BR3_RNA_030_E_fc_df <- RNA_fc_fn("BR3", "BR3_RNA_030_E")
BR3_RNA_060_E_fc_df <- RNA_fc_fn("BR3", "BR3_RNA_060_E")
BR3_RNA_180_E_fc_df <- RNA_fc_fn("BR3", "BR3_RNA_180_E")

# BR4
BR4_RNA_000_E_fc_df <- RNA_fc_fn("BR4", "BR4_RNA_000_E")
BR4_RNA_005_E_fc_df <- RNA_fc_fn("BR4", "BR4_RNA_005_E")
BR4_RNA_015_E_fc_df <- RNA_fc_fn("BR4", "BR4_RNA_015_E")
BR4_RNA_030_E_fc_df <- RNA_fc_fn("BR4", "BR4_RNA_030_E")
BR4_RNA_060_E_fc_df <- RNA_fc_fn("BR4", "BR4_RNA_060_E")
BR4_RNA_180_E_fc_df <- RNA_fc_fn("BR4", "BR4_RNA_180_E")

# Extract read counts, combine, and reformat
RNA_fc_df <- 
  # BR2
  BR2_RNA_000_E_fc_df %>%
  full_join(BR2_RNA_005_E_fc_df, by = "gene_id") %>%
  full_join(BR2_RNA_015_E_fc_df, by = "gene_id") %>%
  full_join(BR2_RNA_030_E_fc_df, by = "gene_id") %>%
  full_join(BR2_RNA_060_E_fc_df, by = "gene_id") %>%
  full_join(BR2_RNA_180_E_fc_df, by = "gene_id") %>%
  # BR3
  full_join(BR3_RNA_000_E_fc_df, by = "gene_id") %>%
  full_join(BR3_RNA_005_E_fc_df, by = "gene_id") %>%
  full_join(BR3_RNA_015_E_fc_df, by = "gene_id") %>%
  full_join(BR3_RNA_030_E_fc_df, by = "gene_id") %>%
  full_join(BR3_RNA_060_E_fc_df, by = "gene_id") %>%
  full_join(BR3_RNA_180_E_fc_df, by = "gene_id") %>%
  # BR4
  full_join(BR4_RNA_000_E_fc_df, by = "gene_id") %>%
  full_join(BR4_RNA_005_E_fc_df, by = "gene_id") %>%
  full_join(BR4_RNA_015_E_fc_df, by = "gene_id") %>%
  full_join(BR4_RNA_030_E_fc_df, by = "gene_id") %>%
  full_join(BR4_RNA_060_E_fc_df, by = "gene_id") %>%
  full_join(BR4_RNA_180_E_fc_df, by = "gene_id") %>%
  column_to_rownames("gene_id")

# Subset featurecounts data to non-p3 genes
RNA_fc_Np3_df <-
  RNA_fc_df %>%
  filter(rownames(RNA_fc_df) %in% Hg_HFTV1_annot_Np3_df$gene_id)

# Subset featurecounts data to p3 genes.
RNA_fc_p3_df <-
  RNA_fc_df %>%
  filter(rownames(RNA_fc_df) %in% Hg_annot_p3_df$gene_id) %>%
  # Remove biological replicate that was missing p3.
  select(!contains("BR2"))


# Create non-p3 metadata
RNA_Np3_meta_df <- 
  data.frame(
    col = colnames(RNA_fc_Np3_df),
    # Time
    time = 
      
      factor(
        
        rep(
          c("000_PI","005_PI","015_PI","030_PI","060_PI","180_PI"), times = 3),
        
        levels = c("000_PI","005_PI","015_PI","030_PI","060_PI","180_PI")),
    
    # Biological replicate
    biorep = 
      factor(c("BR2", "BR2", "BR2", "BR2", "BR2", "BR2", 
               "BR3", "BR3", "BR3", "BR3", "BR3", "BR3",
               "BR4", "BR4", "BR4", "BR4", "BR4", "BR4"),
             levels = c("BR2", "BR3", "BR4")),
    
    # Batch (effectively the same as biorep)
    batch = 
      factor(c("batch1", "batch1", "batch1", "batch1", "batch1", "batch1", 
               "batch2", "batch2", "batch2", "batch2", "batch2", "batch2", 
               "batch3", "batch3", "batch3", "batch3", "batch3", "batch3"),
             levels = c("batch1", "batch2", "batch3")),
    
    # Experimental/infected ("E") vs. control/mock-infected (C); for now, all
    # samples are experimental only.
    infection = 
      factor(rep(c("E"), times = 18)),
    # Sample type (RNA vs RBO)
    sample_type = 
      factor(rep(c("RNA"), times = 18))) %>%
  column_to_rownames("col")

# Create p3 metadata
RNA_p3_meta_df <- 
  data.frame(
    col = colnames(RNA_fc_p3_df),
    # Time (nominal)
    # https://www.biostars.org/p/9499264/
    time = 
      
      factor(
        
        rep(
          c("000_PI","005_PI","015_PI","030_PI","060_PI","180_PI"), times = 2),
        
        levels = c("000_PI","005_PI","015_PI","030_PI","060_PI","180_PI")),
    
    # Biological replicate
    biorep = 
      factor(c("BR3", "BR3", "BR3", "BR3", "BR3", "BR3",
               "BR4", "BR4", "BR4", "BR4", "BR4", "BR4"),
             levels = c("BR2", "BR3", "BR4")),
    
    # Batch (effectively the same as biorep)
    batch = 
      factor(c("batch2", "batch2", "batch2", "batch2", "batch2", "batch2", 
               "batch3", "batch3", "batch3", "batch3", "batch3", "batch3"),
             levels = c("batch2", "batch3")),
    
    # Experimental/infected ("E") vs. control/mock-infected (C); for now, all
    # samples are experimental only.
    infection = 
      factor(rep(c("E"), times = 12)),
    # Sample type (RNA vs RBO)
    sample_type = 
      factor(rep(c("RNA"), times = 12))) %>%
  column_to_rownames("col")

# === RBO ===

RBO_fc_fn <- function(rep, sample) {
  
  dataframe <- 
    # Read in featurecounts data
    read_tsv(paste(
      "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
      "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/d_FEATURECOUNTS/",
      sample, "_fc.txt", sep=""),
             # Skip first line (record of featurecounts command)
             skip=1, 
             # Do not show column type
             show_col_types = FALSE) %>%
    # Rename columns 
    dplyr::rename(
      # Column 1: gene ID
      gene_id = 1,
      # Column 7: sample-specific feature counts
      {{sample}}:=7) %>%
    # Select only the gene_id and read counts columns for further use. 
    dplyr::select(c(1,7)) %>% 
    # Round read count values to nearest integer (for featurecount data
    # generated using --fraction option for reads that map to overlapping
    # features. 
    dplyr::mutate(across(where(is.numeric), round, 0)) %>%
    # Subset to unique entries
    unique()
  
  return(dataframe)
  
}

# BR2 (account for 015 <-> 060)
BR2_RBO_000_E_fc_df <- RBO_fc_fn("BR2", "BR2_RBO_000_E_R1")
BR2_RBO_005_E_fc_df <- RBO_fc_fn("BR2", "BR2_RBO_005_E_R1")
BR2_RBO_015_E_fc_df <- RBO_fc_fn("BR2", "BR2_RBO_060_E_R1") %>%
  dplyr::rename(BR2_RBO_015_E_R1 = BR2_RBO_060_E_R1)
BR2_RBO_060_E_fc_df <- RBO_fc_fn("BR2", "BR2_RBO_015_E_R1") %>%
  dplyr::rename(BR2_RBO_060_E_R1 = BR2_RBO_015_E_R1)
BR2_RBO_180_E_fc_df <- RBO_fc_fn("BR2", "BR2_RBO_180_E_R1")

# BR3 (account for swaps among 005, 030, and 015)
BR3_RBO_000_E_fc_df <- RBO_fc_fn("BR3", "BR3_RBO_000_E_R1")
BR3_RBO_005_E_fc_df <- RBO_fc_fn("BR3", "BR3_RBO_030_E_R1") %>%
  dplyr::rename(BR3_RBO_005_E_R1 = BR3_RBO_030_E_R1)
BR3_RBO_015_E_fc_df <- RBO_fc_fn("BR3", "BR3_RBO_005_E_R1") %>%
  dplyr::rename(BR3_RBO_015_E_R1 = BR3_RBO_005_E_R1)
BR3_RBO_030_E_fc_df <- RBO_fc_fn("BR3", "BR3_RBO_015_E_R1") %>%
  dplyr::rename(BR3_RBO_030_E_R1 = BR3_RBO_015_E_R1)
BR3_RBO_180_E_fc_df <- RBO_fc_fn("BR3", "BR3_RBO_180_E_R1")

# BR4 (account for swaps among 060 <-> 030)
BR4_RBO_000_E_fc_df <- RBO_fc_fn("BR4", "BR4_RBO_000_E_R1")
BR4_RBO_005_E_fc_df <- RBO_fc_fn("BR4", "BR4_RBO_005_E_R1")
BR4_RBO_015_E_fc_df <- RBO_fc_fn("BR4", "BR4_RBO_015_E_R1")
BR4_RBO_030_E_fc_df <- RBO_fc_fn("BR4", "BR4_RBO_060_E_R1") %>%
  dplyr::rename(BR4_RBO_030_E_R1 = BR4_RBO_060_E_R1)
BR4_RBO_060_E_fc_df <- RBO_fc_fn("BR4", "BR4_RBO_030_E_R1") %>%
  dplyr::rename(BR4_RBO_060_E_R1 = BR4_RBO_030_E_R1)
BR4_RBO_180_E_fc_df <- RBO_fc_fn("BR4", "BR4_RBO_180_E_R1")


# Extract read counts, combine, and reformat
RBO_fc_df <-
  # BR2
  BR2_RBO_000_E_fc_df %>%
  full_join(BR2_RBO_005_E_fc_df, by = "gene_id") %>%
  full_join(BR2_RBO_015_E_fc_df, by = "gene_id") %>%
  full_join(BR2_RBO_060_E_fc_df, by = "gene_id") %>%
  full_join(BR2_RBO_180_E_fc_df, by = "gene_id") %>%
  # BR3
  full_join(BR3_RBO_000_E_fc_df, by = "gene_id") %>%
  full_join(BR3_RBO_005_E_fc_df, by = "gene_id") %>%
  full_join(BR3_RBO_015_E_fc_df, by = "gene_id") %>%
  full_join(BR3_RBO_030_E_fc_df, by = "gene_id") %>%
  full_join(BR3_RBO_180_E_fc_df, by = "gene_id") %>%
  # BR4
  full_join(BR4_RBO_000_E_fc_df, by = "gene_id") %>%
  full_join(BR4_RBO_005_E_fc_df, by = "gene_id") %>%
  full_join(BR4_RBO_015_E_fc_df, by = "gene_id") %>%
  full_join(BR4_RBO_030_E_fc_df, by = "gene_id") %>%
  full_join(BR4_RBO_060_E_fc_df, by = "gene_id") %>%
  full_join(BR4_RBO_180_E_fc_df, by = "gene_id") %>%

  column_to_rownames("gene_id")

# Subset non-p3 genetic elements
RBO_fc_Np3_df <-
  RBO_fc_df %>%
  filter(rownames(RBO_fc_df) %in% Hg_HFTV1_annot_Np3_df$gene_id)

# Create non-p3 metadata
RBO_Np3_meta_df <- 
  data.frame(
    col = colnames(RBO_fc_Np3_df),
    # Time
    time = 
      factor(c(
        # BR2
        "000_PI", "005_PI", "015_PI", "060_PI", "180_PI",
        # BR3
        "000_PI", "005_PI", "015_PI", "030_PI", "180_PI",
        # BR4
        "000_PI", "005_PI", "015_PI", "030_PI", "060_PI", "180_PI"),
        
        # Define time points as distinct levels of time factor
        levels = c(
          "000_PI", "005_PI", "015_PI", "030_PI", 
          "060_PI", "180_PI")),
    
    # Biological replicate
    biorep = 
      factor(c("BR2", "BR2", "BR2", "BR2", "BR2", 
               "BR3", "BR3", "BR3", "BR3", "BR3",
               "BR4", "BR4", "BR4", "BR4", "BR4", "BR4"),
             levels = c("BR2", "BR3", "BR4")),
    
    # Batch (effectively the same as biological replicate)
    batch = 
      factor(c("batch1", "batch1", "batch1", "batch1", "batch1", 
               "batch2", "batch2", "batch2", "batch2", "batch2",
               "batch3", "batch3", "batch3", "batch3", "batch3", "batch3"),
             levels = c("batch1", "batch2", "batch3")),
    
    # Experimental/infected ("E") vs. control/mock-infected (C); for now, all
    # samples are experimental only.
    infection = 
      factor(rep(c("E"), times = 16)),
    # Sample type (RNA vs RBO)
    sample_type = 
      factor(rep(c("RBO"), times = 16))) %>%
  column_to_rownames("col")

```

# === DEA | ISO | DDS ========================================================
```{r}
dds_generate_filter_fn <- function(count_data, col_data) {
  
  # Construct DESeqDataSet (DDS)
  dds <- DESeqDataSetFromMatrix(
    countData = count_data,
    colData = col_data,
    design = ~ time)
  # batch cannot be incorporated into experimental design when 200 min 
  # PI samples are included. 
    # design = ~ batch + time)
  
  # Define values for pre-filtering. 
  
  # smallestCountSize: Minimum number of counts per gene
  smallestCountSize <- 50
  # smallestGroupSize: Minimum number of groups in which gene is expressed.
  # If there are not discrete groups, use the minimal number of samples where
  # non-zero counts would be considered interesting. 
  smallestGroupSize <- 2
  
  # Filter DDS
  keep <- rowSums(counts(dds) >= smallestCountSize) >=
    smallestGroupSize
  dds_filter <- dds[keep,]
  
  return(dds_filter)
}

# === RNA ===

RNA_Np3_dds <- 
  dds_generate_filter_fn(RNA_fc_Np3_df, RNA_Np3_meta_df)
RNA_Np3_dds_vals <- assay(RNA_Np3_dds)
RNA_p3_dds <- 
  dds_generate_filter_fn(RNA_fc_p3_df, RNA_p3_meta_df)
RNA_p3_dds_vals <- assay(RNA_p3_dds)

# === RBO ===
RBO_Np3_dds <- 
  dds_generate_filter_fn(RBO_fc_Np3_df, RBO_Np3_meta_df)
RBO_Np3_dds_vals <- assay(RBO_Np3_dds)
# RBO_p3_dds <- 
#   dds_generate_filter_fn(RBO_fc_p3_df, RBO_p3_meta_df)
# RBO_p3_dds_vals <- assay(RBO_p3_dds)

```

# === TEA | SFN ==============================================================
```{r}
# === RNA ===

# Np3
RNA_Np3_sfn <- estimateSizeFactors(RNA_Np3_dds)
RNA_Np3_sfn_df <- as.data.frame(assay(RNA_Np3_sfn)) %>%
  rownames_to_column("gene_id")
# p3
RNA_p3_sfn <- estimateSizeFactors(RNA_p3_dds)
RNA_p3_sfn_df <- as.data.frame(assay(RNA_p3_sfn)) %>%
  rownames_to_column("gene_id")


# === RBO ===

# Np3
RBO_Np3_sfn <- estimateSizeFactors(RBO_Np3_dds)
RBO_Np3_sfn_df <- as.data.frame(assay(RBO_Np3_sfn)) %>%
  rownames_to_column("gene_id")
# p3
# RBO_p3_sfn <- estimateSizeFactors(RBO_p3_dds)
# RBO_p3_sfn_df <- as.data.frame(assay(RBO_p3_sfn)) %>%
#   rownames_to_column("gene_id")
```

# === TEA | ISO | TRANSFORM ==================================================
```{r}
# === RNA ===

# Np3
RNA_Np3_rld <- rlog(RNA_Np3_dds, blind = FALSE)
RNA_Np3_rld_df <- as.data.frame(assay(RNA_Np3_rld)) %>%
  rownames_to_column("gene_id")
# p3
RNA_p3_rld <- rlog(RNA_p3_dds, blind = FALSE)
RNA_p3_rld_df <- as.data.frame(assay(RNA_p3_rld)) %>%
  rownames_to_column("gene_id")


# === RBO ===

# Np3
RBO_Np3_rld <- rlog(RBO_Np3_dds, blind = FALSE)
RBO_Np3_rld_df <- as.data.frame(assay(RBO_Np3_rld)) %>%
  rownames_to_column("gene_id")
# p3
# RBO_p3_rld <- rlog(RBO_p3_dds, blind = FALSE)
# RBO_p3_rld_df <- as.data.frame(assay(RBO_p3_rld)) %>%
#   rownames_to_column("gene_id")


```

# === TEA | ISO | TRANSFORM | SUMMARY STATISTICS | FUN =======================
```{r}
transf_summary_fn <- 
  function(transf_df) {
    
    
    # === PRELIMINARIES ===
    
    # Based on input dataframe name, construct vector of "arguments" that
    # describe some features of the input dataframe. Assumes transf_df name
    # structure of <sample_type>_<genetic_elements>_<transf_type>_...
    df_args <- str_split(deparse(substitute(transf_df)), "_", simplify = TRUE)
    
    # Sample type (RNA vs. RBO)
    sample_type <- df_args[1]
    
    # Genetic elements (Np3 vs p3)
    genetic_elements <- df_args[2]
    
    # Transformation type (rlog [rld], variance-stabilized[vsd], size-factor-
    # normalized [sfnorm], etc.)
    transf_type <- df_args[3]
    
    transf_df <-
      transf_df %>%
      # Convert gene IDs to rownames so that they are preserved even after
      # selection of specific columns during various subsetting steps below. 
      column_to_rownames("gene_id")
    
    # Account for missing samples (no missing samples in RNA, but two
    # missing across all RBO replicates).
    
    if (sample_type == "RNA") {
      
      col_names <- c("gene_id", 0, 5, 15, 30, 60, 180)
      
      BR2_col_names <- col_names
      BR3_col_names <- col_names
      BR4_col_names <- col_names
      
    } else if (sample_type == "RBO") {
      
      # Account for missing data for pHGLR3
      if (genetic_elements == "p3") {
        
        # Redefine column names to rename subset columns for graphing later
        # (no 60 min sample in RBO data; there was only one replicate after
        # BR2 was removed, so I took it out entirely).
        
        col_names <- c("gene_id", 0, 5, 15, 30, 180)
        
        BR3_col_names <- c("gene_id", 0, 5, 15, 30, 180)
        BR4_col_names <- c("gene_id", 0, 5, 15, 30, 180)
        
      } else if (genetic_elements == "Np3") {
        
        col_names <- c("gene_id", 0, 5, 15, 30, 60, 180)
        
        BR2_col_names <- c("gene_id", 0, 5, 15, 60, 180)
        BR3_col_names <- c("gene_id", 0, 5, 15, 30, 180)
        BR4_col_names <- c("gene_id", 0, 5, 15, 30, 60, 180)
      }
    }
    
    
    # === SUBSET TRANSFORMED EXPRESSION VALUES BY BIOLOGICAL REPLICATE ===

    # These or mean values can be used for DP-GP-Cluster.

    BR2_transf_df <- transf_df %>% select(contains("BR2")) %>%
      rownames_to_column("gene_id")
    # Only try to rename BR2 column names for non-p3 genetic elements. BR2
    # data for p3 do not exist.
    if (genetic_elements == "Np3") {
      colnames(BR2_transf_df) <- BR2_col_names
    }
    
    BR3_transf_df <- transf_df %>% select(contains("BR3")) %>%
      rownames_to_column("gene_id")
    colnames(BR3_transf_df) <- BR3_col_names
    
    BR4_transf_df <- transf_df %>% select(contains("BR4")) %>%
      rownames_to_column("gene_id")
    colnames(BR4_transf_df) <- BR4_col_names
    



    # === CALCULATE SUMMARY STATISTICS ===

    # Calculate mean expression value for each row (gene) and at each time
    # point ("000", "005", "015",...) across biological replicates.
    transf_000_E_summary_df <- 
      transf_df %>%
      # Select columns for specific time point / condition.
      select(contains("000_E")) %>%
      mutate(
        # Calculate row-wise mean across all selected columns.
        mean_000_E = rowMeans(.),
        # Calculate row-wise standard deviation across all selected columns.
        stdev_000_E = rowSds(as.matrix(.)))
    # Do the same as above for all other time points / conditions.
    transf_005_E_summary_df <- transf_df %>% select(contains("005_E")) %>%
      mutate(mean_005_E = rowMeans(.), stdev_005_E = rowSds(as.matrix(.)))
    transf_015_E_summary_df <- transf_df %>% select(contains("015_E")) %>%
      mutate(mean_015_E = rowMeans(.), stdev_015_E = rowSds(as.matrix(.)))
    transf_030_E_summary_df <- transf_df %>% select(contains("030_E")) %>%
      mutate(mean_030_E = rowMeans(.), stdev_030_E = rowSds(as.matrix(.)))
    
    # Conditionally calculate summary stats for 060 TP to exclude RBO p3
    if (genetic_elements == "Np3") {
      transf_060_E_summary_df <- transf_df %>% select(contains("060_E")) %>%
        mutate(mean_060_E = rowMeans(.), stdev_060_E = rowSds(as.matrix(.)))
    } else if (genetic_elements == "p3") {
      if (sample_type == "RNA") {
        transf_060_E_summary_df <- transf_df %>% select(contains("060_E")) %>%
          mutate(mean_060_E = rowMeans(.), stdev_060_E = rowSds(as.matrix(.)))
      }
    }
    
    transf_180_E_summary_df <- transf_df %>% select(contains("180_E")) %>%
      mutate(mean_180_E = rowMeans(.), stdev_180_E = rowSds(as.matrix(.)))

    # Combine all mean values into a single dataframe
    transf_summary_df <-
      merge(transf_000_E_summary_df, transf_005_E_summary_df,
            by = "row.names") %>%
      column_to_rownames("Row.names")
    transf_summary_df <-
      merge(transf_015_E_summary_df, transf_summary_df, by = "row.names") %>%
      column_to_rownames("Row.names")
    transf_summary_df <-
      merge(transf_030_E_summary_df, transf_summary_df, by = "row.names") %>%
      column_to_rownames("Row.names")
    # Include 060 time point for RNA regardless of genetic elements
    if (sample_type == "RNA") {
      transf_summary_df <-
        merge(transf_060_E_summary_df, transf_summary_df, by = "row.names") %>%
        column_to_rownames("Row.names")
      # Include 060 time point for RBO only if dealing with non-p3 genetic
      # elements.
    } else if (sample_type == "RBO") {
      
      if (genetic_elements == "Np3") {
        transf_summary_df <-
          merge(transf_060_E_summary_df, transf_summary_df, by = "row.names") %>%
          column_to_rownames("Row.names")
      }
    }
    
    transf_summary_df <-
      merge(transf_180_E_summary_df, transf_summary_df, by = "row.names") %>%
      column_to_rownames("Row.names")


    # === REFORMAT AND SUBSET DATA ===

    transf_summary_df <-
      transf_summary_df %>%
      # Order columns alphabetically
      select(order(colnames(transf_summary_df))) %>%
      # Convert row names to column
      rownames_to_column("gene_id")

    # Subset to means and gene IDs only
    transf_means_df <-
      transf_summary_df %>%
      select(contains(c("gene_id", "mean")))

    # Subset to stdev and gene IDs only
    transf_stdev_df <-
      transf_summary_df %>%
      select(contains(c("gene_id", "stdev")))

    # Rename column names for graphing, adjusting accordingly for RBO p3
    colnames(transf_means_df) <- col_names
    # if (sample_type == "RBO") {
    #   if (genetic_elements == "p3") {
    #     colnames(transf_means_df) <- col_names
    #     colnames(transf_stdev_df) <- col_names
    #   } else if (genetic_elements == "Np3") {
    #     colnames(transf_means_df) <- col_names
    #     colnames(transf_stdev_df) <- col_names
    #   }
    # } else if (sample_type == "RNA") {
    #   colnames(transf_means_df) <- col_names
    #   colnames(transf_stdev_df) <- col_names
    # }


    # === SCALE MEAN VALUES ===

    # Scaled data can be used in R to visualize DP-GP-clusted genes.
    # Convert gene IDs to row names so they are not affected during
    # transposition and scaling.
    transf_means_to_transpose_df <-
      transf_means_df %>%
      column_to_rownames("gene_id")
    # Transpose data for scaling
    transf_means_t_df <- as.data.frame(t(transf_means_to_transpose_df))
    # Scale data (for each gene, mean-center and then divide by stdev)
    transf_means_t_scaled_df <- as.data.frame(scale(transf_means_t_df))
    # Transpose scaled data again to return it to the original shape.
    transf_means_scaled_df <- as.data.frame(t(transf_means_t_scaled_df))
    # Convert gene IDs back to column for downstream applications.
    transf_means_scaled_df <-
      transf_means_scaled_df %>%
      rownames_to_column("gene_id")


    # # === EXPORTS ===
    # 
    # # Individual biological replicates
    # write_csv(
    #   BR2_transf_df,
    #   paste("../c_ANA/c_DE_ANALYSIS/a_ISO/c_RLD/a_DATA/",
    #         genetic_elements, "/", sample_type, "_BR2_", genetic_elements, "_",
    #         transf_type, ".csv", sep = ""))
    # write_csv(
    #   BR3_transf_df,
    #   paste("../c_ANA/c_DE_ANALYSIS/a_ISO/c_RLD/a_DATA/",
    #         genetic_elements, "/", sample_type, "_BR3_", genetic_elements, "_",
    #         transf_type, ".csv", sep = ""))
    # write_csv(
    #   BR4_transf_df,
    #   paste("../c_ANA/c_DE_ANALYSIS/a_ISO/c_RLD/a_DATA/",
    #         genetic_elements, "/", sample_type, "_BR4_", genetic_elements, "_",
    #         transf_type, ".csv", sep = ""))
    # 
    # if (sample_type == "RBO") {
    #   
    #   write_csv(
    #     BR5_transf_df,
    #     paste("../c_ANA/c_DE_ANALYSIS/a_ISO/c_RLD/a_DATA/",
    #           genetic_elements, "/", sample_type, "_BR5_", genetic_elements, "_",
    #           transf_type, ".csv", sep = ""))
    #   write_csv(
    #     BR6_transf_df,
    #     paste("../c_ANA/c_DE_ANALYSIS/a_ISO/c_RLD/a_DATA/",
    #           genetic_elements, "/", sample_type, "_BR6_", genetic_elements, "_",
    #           transf_type, ".csv", sep = ""))
    #   
    # }
    # 
    # # Summary statistics
    # write_csv(
    #   transf_summary_df,
    #   paste("../c_ANA/c_DE_ANALYSIS/a_ISO/c_RLD/b_SUMMARY_STATS/",
    #         genetic_elements, "/", sample_type, "_", genetic_elements, "_",
    #         transf_type, "_summary.csv", sep = ""))
    # # Means only
    # write_csv(
    #   transf_means_df,
    #   paste("../c_ANA/c_DE_ANALYSIS/a_ISO/c_RLD/b_SUMMARY_STATS/",
    #         genetic_elements, "/", sample_type, "_", genetic_elements, "_",
    #         transf_type, "_means.csv", sep = ""))
    # # Stdev only
    # write_csv(
    #   transf_stdev_df,
    #   paste("../c_ANA/c_DE_ANALYSIS/a_ISO/c_RLD/b_SUMMARY_STATS/",
    #         genetic_elements, "/", sample_type, "_", genetic_elements, "_",
    #         transf_type, "_stdev.csv", sep = ""))
    # # Scaled means only
    # write_csv(
    #   transf_means_scaled_df,
    #   paste("../c_ANA/c_DE_ANALYSIS/a_ISO/c_RLD/b_SUMMARY_STATS/",
    #         genetic_elements, "/", sample_type, "_", genetic_elements, "_",
    #         transf_type, "_means_scaled.csv", sep = ""))


    # === FUNCTION OUTPUT ===

    fn_output <-
      list("transf_summary_df" = transf_summary_df,
           "transf_means_df" = transf_means_df,
           "transf_stdev_df" = transf_stdev_df,
           "transf_means_scaled_df" = transf_means_scaled_df)

    return(fn_output)
}

```

# === TEA | ISO | TRANSFORM | SUMMARY STATISTICS =============================
```{r}

# ===========
# === RNA ===
# ===========

# === Np3 ===

# Summarize data
RNA_Np3_rld_summary <-
  transf_summary_fn(
    transf_df = RNA_Np3_rld_df)

# Break summarization function output into dataframes
RNA_Np3_rld_summary_df <- RNA_Np3_rld_summary$transf_summary_df
RNA_Np3_rld_means_df <- RNA_Np3_rld_summary$transf_means_df
RNA_Np3_rld_stdev_df <- RNA_Np3_rld_summary$transf_stdev_df
RNA_Np3_rld_means_scl_df <- RNA_Np3_rld_summary$transf_means_scaled_df

# === p3 ===

# Summarize data
RNA_p3_rld_summary <-
  transf_summary_fn(
    transf_df = RNA_p3_rld_df)

# Break summarization function output into dataframes
RNA_p3_rld_summary_df <- RNA_p3_rld_summary$transf_summary_df
RNA_p3_rld_means_df <- RNA_p3_rld_summary$transf_means_df
RNA_p3_rld_stdev_df <- RNA_p3_rld_summary$transf_stdev_df
RNA_p3_rld_means_scl_df <- RNA_p3_rld_summary$transf_means_scaled_df

# === AGE ===

# Combine transformed means data for AGE
RNA_AGE_rld_means_df <-
  rbind(RNA_Np3_rld_means_df, RNA_p3_rld_means_df)
# Compile transformed mean scaled expression values for AGE
RNA_AGE_rld_means_scl_df <-
  rbind(RNA_Np3_rld_means_scl_df, RNA_p3_rld_means_scl_df)


# ===========
# === RBO ===
# ===========

# === Np3 ===

# Summarize data
RBO_Np3_rld_summary <-
  transf_summary_fn(
    transf_df = RBO_Np3_rld_df)

# Break summarization function output into dataframes
RBO_Np3_rld_summary_df <- RBO_Np3_rld_summary$transf_summary_df
RBO_Np3_rld_means_df <- RBO_Np3_rld_summary$transf_means_df
RBO_Np3_rld_stdev_df <- RBO_Np3_rld_summary$transf_stdev_df
RBO_Np3_rld_means_scl_df <- RBO_Np3_rld_summary$transf_means_scaled_df

# === p3 ===

# # Summarize data
# RBO_p3_rld_summary <-
#   transf_summary_fn(
#     transf_df = RBO_p3_rld_df)
# 
# # Break summarization function output into dataframes
# RBO_p3_rld_summary_df <- RBO_p3_rld_summary$transf_summary_df
# RBO_p3_rld_means_df <- RBO_p3_rld_summary$transf_means_df
# RBO_p3_rld_stdev_df <- RBO_p3_rld_summary$transf_stdev_df
# RBO_p3_rld_means_scl_df <- RBO_p3_rld_summary$transf_means_scaled_df


```

# === TEA | SFN | SUMMARY STATISTICS =========================================
```{r}
# ===========
# === RNA ===
# ===========

# === Np3 ===

# Summarize data
RNA_Np3_sfn_summary <-
  transf_summary_fn(
    transf_df = RNA_Np3_sfn_df)

# Break summarization function output into dataframes
RNA_Np3_sfn_summary_df <- RNA_Np3_sfn_summary$transf_summary_df
RNA_Np3_sfn_means_df <- RNA_Np3_sfn_summary$transf_means_df
RNA_Np3_sfn_stdev_df <- RNA_Np3_sfn_summary$transf_stdev_df
RNA_Np3_sfn_means_scl_df <- RNA_Np3_sfn_summary$transf_means_scaled_df

# === p3 ===

# Summarize data
RNA_p3_sfn_summary <-
  transf_summary_fn(
    transf_df = RNA_p3_sfn_df)

# Break summarization function output into dataframes
RNA_p3_sfn_summary_df <- RNA_p3_sfn_summary$transf_summary_df
RNA_p3_sfn_means_df <- RNA_p3_sfn_summary$transf_means_df
RNA_p3_sfn_stdev_df <- RNA_p3_sfn_summary$transf_stdev_df
RNA_p3_sfn_means_scl_df <- RNA_p3_sfn_summary$transf_means_scaled_df

# === AGE ===

# Combine transformed means data for AGE
RNA_AGE_sfn_means_df <-
  rbind(RNA_Np3_sfn_means_df, RNA_p3_sfn_means_df)
# Compile transformed mean scaled expression values for AGE
RNA_AGE_sfn_means_scl_df <-
  rbind(RNA_Np3_sfn_means_scl_df, RNA_p3_sfn_means_scl_df)


# ===========
# === RBO ===
# ===========

# === Np3 ===

# Summarize data
RBO_Np3_sfn_summary <-
  transf_summary_fn(
    transf_df = RBO_Np3_sfn_df)

# Break summarization function output into dataframes
RBO_Np3_sfn_summary_df <- RBO_Np3_sfn_summary$transf_summary_df
RBO_Np3_sfn_means_df <- RBO_Np3_sfn_summary$transf_means_df
RBO_Np3_sfn_stdev_df <- RBO_Np3_sfn_summary$transf_stdev_df
RBO_Np3_sfn_means_scl_df <- RBO_Np3_sfn_summary$transf_means_scaled_df

# === p3 ===

# # Summarize data
# RBO_p3_sfn_summary <-
#   transf_summary_fn(
#     transf_df = RBO_p3_sfn_df)
# 
# # Break summarization function output into dataframes
# RBO_p3_sfn_summary_df <- RBO_p3_sfn_summary$transf_summary_df
# RBO_p3_sfn_means_df <- RBO_p3_sfn_summary$transf_means_df
# RBO_p3_sfn_stdev_df <- RBO_p3_sfn_summary$transf_stdev_df
# RBO_p3_sfn_means_scl_df <- RBO_p3_sfn_summary$transf_means_scaled_df
# 
# 
# 
# # Combine transformed means data for AGE
# # RBO rbind must be .fill to account for missing 060 information in p3 data.
# RBO_AGE_sfn_means_df <-
#   plyr::rbind.fill(RBO_Np3_sfn_means_df, RBO_p3_sfn_means_df)
# # Compile transformed mean scaled expression values for AGE
# # RBO rbind must be .fill to account for missing 060 information in p3 data.
# RBO_AGE_sfn_means_scl_df <-
#   plyr::rbind.fill(RBO_Np3_sfn_means_scl_df, RBO_p3_sfn_means_scl_df)
```

# === TEA | SFN | WRANGLE ====================================================
```{r}

# RNA
RNA_AGE_sfn_means_long_df <-
  RNA_AGE_sfn_means_df %>%
  pivot_longer(
    cols = !gene_id,
    names_to = "time_PI",
    values_to = "RNA_sfn")

# RBO
RBO_Np3_sfn_means_long_df <-
  RBO_Np3_sfn_means_df %>%
  pivot_longer(
    cols = !gene_id,
    names_to = "time_PI",
    values_to = "RBO_sfn")

# Combine RNA and RBO; calculate TE
RnR_Np3_sfn_means_long_df <-
  inner_join(
    RNA_AGE_sfn_means_long_df %>%
      select(gene_id, time_PI, RNA_sfn),
    RBO_Np3_sfn_means_long_df %>%
      select(gene_id, time_PI, RBO_sfn),
    by = c("gene_id", "time_PI")) %>%
  mutate(
    RNA_sfn = RNA_sfn + 1,
    RBO_sfn = RBO_sfn + 1,
    TE = RBO_sfn/RNA_sfn,
    RNA_log10 = log10(RNA_sfn),
    RBO_log10 = log10(RBO_sfn),
    TE_log10 = log10(TE)) 

# Add entity information
RnR_Np3_sfn_means_long_df <-
  RnR_Np3_sfn_means_long_df %>% 
  mutate(
    entity = 
      case_when(
        str_detect(gene_id, "M1M17") ~ "HFTV1", TRUE ~ "Hg"))

RnR_Np3_sfn_means_long_df$time_PI <- 
  as.numeric(RnR_Np3_sfn_means_long_df$time_PI)


# L2FC relative to a specific time point
RnR_Np3_sfn_means_long_ref_df <- 
  RnR_Np3_sfn_means_long_df %>% 
  filter(time_PI == 0) %>%
  select(c(gene_id, TE)) %>%
  dplyr::rename(TE_ref = TE)

RnR_Np3_sfn_means_long_L2FC_df <-
  left_join(
    RnR_Np3_sfn_means_long_df,
    RnR_Np3_sfn_means_long_ref_df,
    by = "gene_id") %>%
  mutate(
    TE_log10_noref = log10(TE),
    FC_TE_vs = TE/TE_ref,
    L2FC_TE_vs = log2(FC_TE_vs))

# RnR_sfn_L2FC_clu_fam_df <-
#   left_join(
#     RnR_AGE_sfn_means_long_L2FC_df,
#     RBO_AGE_clu_fam_annot_df %>% select(c(clu_fam, gene_id)),
#     by = "gene_id")


RnR_Np3_sfn_means_longer_df <-
  RnR_Np3_sfn_means_long_df %>%
  pivot_longer(
    cols = !c(gene_id, time_PI, entity),
    names_to = "category",
    values_to = "expr_value") %>%
  # Remove 188 rows of p3 genes with NA values
  filter(!is.na(expr_value))

RnR_Np3_sfn_means_longer_df$category <-
  factor(
    RnR_Np3_sfn_means_longer_df$category,
    levels = c(
      "RNA_sfn",
      "RBO_sfn",
      "TE",
      "RNA_log10",
      "RBO_log10",
      "TE_log10"
    ))
```

# === TEA | TE CALC ==========================================================
```{r}
# RNA
RNA_AGE_sfn_means_long_df <-
  RNA_AGE_sfn_means_df %>%
  pivot_longer(
    cols = !gene_id,
    names_to = "time_PI",
    values_to = "RNA_sfn")

# RBO
RBO_Np3_sfn_means_long_df <-
  RBO_Np3_sfn_means_df %>%
  pivot_longer(
    cols = !gene_id,
    names_to = "time_PI",
    values_to = "RBO_sfn")

# Combine RNA and RBO
TE_df <-
  inner_join(
    RNA_AGE_sfn_means_long_df %>%
      select(gene_id, time_PI, RNA_sfn),
    RBO_Np3_sfn_means_long_df %>%
      select(gene_id, time_PI, RBO_sfn),
    by = c("gene_id", "time_PI")) %>%
  mutate(
    # Add entity information
    entity = 
      case_when(
        str_detect(gene_id, "M1M17") ~ "HFTV1", TRUE ~ "Hg"),
    # Pseudocounts prior to TE calculation
    RNA_sfn = RNA_sfn + 1,
    RBO_sfn = RBO_sfn + 1,
    # Calculate TE
    TE = RBO_sfn/RNA_sfn,
    RNA_sfn_log2 = log2(RNA_sfn),
    RBO_sfn_log2 = log2(RBO_sfn),
    TE_log2 = log2(TE)) 

TE_df$time_PI <- 
  as.numeric(TE_df$time_PI)
```

# === TEA | TE CALC | EXPORT =================================================
```{r}
# write_csv(
#   TE_df %>% filter(grepl("gp17", gene_id)),
#   paste(
#     "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/",
#     "H_GIBBONSII_HFTV1/RNA_RBO/c_ANA/b_REANNOTATION/MNL/",
#     "i_GOIS/gp17_two_TSSs/f_TE/TE_splitgp17_HFTV1_gp17.csv", sep = ""
#   )
# )
```

# === TEA | TE CALC | IMPORT (*) =============================================
```{r}
TE_splitgp17_HFTV1_df <-
  read_csv(
    paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/",
    "RNA_RBO/c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/f_TE/a_DATA/",
    "TE_splitgp17_HFTV1.csv", sep = ""),
    show_col_types = FALSE)
```

# === TEA | VIS ==============================================================
```{r}

test_plot <-
  
  ggplot(
    TE_splitgp17_HFTV1_df %>% 
      filter(grepl("gp17", gene_id)),
    aes(
      x = time_PI,
      y = TE,
      group = gene_id,
      color = gene_id,
      linetype = gene_id)) +
  
  geom_line(
    linewidth = 1, 
    alpha = 1,
    aes(color = gene_id)) +
  geom_point(
    size = 2.5, 
    shape = 21,
    color = "black",
    stroke = 1,
    aes(fill = gene_id)) +
  
  scale_color_manual(
    values = c(
      M1M17_gp17_CTERM = "indianred3",
      M1M17_gp17_NTERM = "steelblue2")) +

  scale_fill_manual(
    values = c(
      M1M17_gp17_CTERM = "indianred3",
      M1M17_gp17_NTERM = "steelblue2")) +
  
  scale_linetype_manual(
    values = c(
      M1M17_gp17_CTERM = "solid",
      M1M17_gp17_NTERM = "dashed")) +
  
  scale_y_continuous(limits = c(0, 90), breaks = seq(0, 100, by = 10)) +
  
  # ylim(0, 10) +
  
  xlab("Time post-infection (minutes)") +
  # ylab("log2( translation efficiency )") +
  
  # facet_wrap(
  #   ~category,
  #   scales = "free_y"
  #   ) +

  theme_classic() +
  theme(
    plot.title=element_text(size=18, face="bold", hjust = 0.5),

    axis.title = element_text(size = 16, face="bold"),
    axis.title.x = element_text(margin=margin(t=10)),
    axis.title.y = element_text(margin=margin(r=10)),

    axis.text = element_text(size = 14, face="bold", color="black"),

    axis.ticks = element_line(linewidth = 0.75, color = "black"),
    axis.ticks.length=unit(0.25, "cm"),

    axis.line=element_line(linewidth = 0, color = "black"),

    legend.text = element_text(face="bold", size=12),
    # No legend
    legend.position="none",

    panel.background = element_rect(fill = "white"),
    panel.border =
      element_rect(colour = "black", fill=NA, linewidth=1.75),
    panel.spacing = unit(7, "lines"),

    strip.text.x =
      element_text(margin = margin(2, 0, 2, 0), size = 14, face="bold"),
    strip.background = element_blank(),
    panel.grid.major.y = element_line(color = "grey80"),
    panel.grid.minor.y = element_line(color = "grey80")
  )


ggsave(
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/RNA_RBO/",
    "c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/f_TE/b_VIS/",
    "TE_by_gp17_region.png", 
    sep = ""),
  plot = test_plot,
  width = 8,
  height = 14,
  units = "cm", 
  dpi = 300)

ggsave(
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/RNA_RBO/",
    "c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/f_TE/b_VIS/",
    "TE_by_gp17_region.svg", 
    sep = ""),
  plot = test_plot,
  width = 8,
  height = 14,
  units = "cm", 
  dpi = 300)
```

# === TEA | VIS | LOG ========================================================
```{r}

test_plot <-
  
  ggplot(
    TE_splitgp17_HFTV1_df %>% 
      filter(grepl("gp17", gene_id)),
    aes(
      x = time_PI,
      y = log2(TE),
      group = gene_id,
      color = gene_id,
      linetype = gene_id)) +
  
  geom_line(
    linewidth = 1, 
    alpha = 1,
    aes(color = gene_id)) +
  geom_point(
    size = 2.5, 
    shape = 21,
    color = "black",
    stroke = 1,
    aes(fill = gene_id)) +
  
  scale_color_manual(
    values = c(
      M1M17_gp17_CTERM = "indianred3",
      M1M17_gp17_NTERM = "steelblue2")) +

  scale_fill_manual(
    values = c(
      M1M17_gp17_CTERM = "indianred3",
      M1M17_gp17_NTERM = "steelblue2")) +
  
  scale_linetype_manual(
    values = c(
      M1M17_gp17_CTERM = "solid",
      M1M17_gp17_NTERM = "dashed")) +
  
  # scale_y_continuous(limits = c(0, 90), breaks = seq(0, 100, by = 10)) +
  
  ylim(0, 8) +
  
  xlab("Time post-infection\n(minutes)") +
  # ylab("log2( translation efficiency )") +
  
  # facet_wrap(
  #   ~category,
  #   scales = "free_y"
  #   ) +

  theme_classic() +
  theme(
    plot.title=element_text(size=18, face="bold", hjust = 0.5),

    axis.title = element_text(size = 16, face="bold"),
    axis.title.x = element_text(margin=margin(t=10)),
    axis.title.y = element_text(margin=margin(r=10)),

    axis.text = element_text(size = 14, face="bold", color="black"),

    axis.ticks = element_line(linewidth = 0.75, color = "black"),
    axis.ticks.length=unit(0.25, "cm"),

    axis.line=element_line(linewidth = 0, color = "black"),

    legend.text = element_text(face="bold", size=12),
    # No legend
    legend.position="none",

    panel.background = element_rect(fill = "white"),
    panel.border =
      element_rect(colour = "black", fill=NA, linewidth=1.75),
    panel.spacing = unit(7, "lines"),

    strip.text.x =
      element_text(margin = margin(2, 0, 2, 0), size = 14, face="bold"),
    strip.background = element_blank(),
    panel.grid.major.y = element_line(color = "grey80"),
    panel.grid.minor.y = element_line(color = "grey80")
  )


ggsave(
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/RNA_RBO/",
    "c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/f_TE/b_VIS/",
    "TE_by_gp17_region_log2.png", 
    sep = ""),
  plot = test_plot,
  width = 8,
  height = 9,
  units = "cm", 
  dpi = 300)

ggsave(
  paste(
    "/Users/mi/DIRUGGIERO_LAB/PROJECTS_AND_RESULTS/H_GIBBONSII_HFTV1/RNA_RBO/",
    "c_ANA/b_REANNOTATION/MNL/i_GOIS/gp17_two_TSSs/f_TE/b_VIS/",
    "TE_by_gp17_region_log2.svg", 
    sep = ""),
  plot = test_plot,
  width = 8,
  height = 9,
  units = "cm", 
  dpi = 300)
```










